### 2025-04-11 code/analysis-code/statistical-analysis-v2-slim-models.qmd
Part 4 additions

---
title: "Processing CDC hypertension and CDV Mortality disease data from adults in the USA exploratory analysis script"
format: html
editor: visual
---

##Setup

```{r setup}
#load needed packages. make sure they are installed.
#install.packages("gt")
#install.packages("emmeans")
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidyverse)
library(tidymodels)
library(gt)
library(stringr)
library(patchwork)
library(emmeans)
library(sf) #for geospatial analysis
library(spdep)
library(spatialreg)
library(vip) # For variable importance plots
library(dplyr)
library(tidytext)
library(forcats)

```

```{r}
#path to data
data_location <- here::here("data","processed-data","eda_outputs.rds")

#load data. 
final_processed_data <- readRDS(data_location)

# Extract key datasets
processed_data_no_percent <- final_processed_data[["processed_data_no_percent"]]
county_race_sex <- final_processed_data[["county_race_sex"]]
county_race <- final_processed_data[["county_race"]]
county_sex <- final_processed_data[["county_sex"]]
county_overall <- final_processed_data[["county_overall"]]

```

#Data fitting/statistical analysis

##Aim 1: Temporal Trends by Age Group

We evaluated national trends in mortality rates over time using multiple modeling approaches. These included a simple linear regression, generalized linear models (GLMs) with race, sex, and age group predictors, and interaction models to assess differential trends. All models assumed a Gaussian distribution and were fit using complete national data.

####1.1 Linear Model - Overall Mortality over time

Fit a basic linear regression to estimate overall trend in mortality over time

```{r}
# Checking the relationship between mortality and year
lm_model_year <- lm(mortalities ~ Year, data = county_overall)
lm_results_year <- broom::tidy(lm_model_year)

# Display and save the regression results
print(lm_results_year)
saveRDS(lm_results_year, file = here::here("products", "manuscript", "supplement", "lm_results_year.rds"))

# Plot linear regression results
lm_plot_year <- ggplot(county_overall, aes(x = Year, y = mortalities)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  theme_minimal() +
  labs(title = "Linear Regression: Mortality Rates Over Time", x = "Year", y = "Mortality Rate")

print(lm_plot_year)
ggsave(here::here("products", "manuscript", "supplement", "lm_plot_year.png"), plot = lm_plot_year)

```

This model provides a first look at overall temporal trends in mortality rates. The flat trend line suggests that overall mortality rates changed minimally from 2000 to 2019, though high variability exists.

#####1.2 GLM: Race and Sex as predictors
GLM with race and sex as additive predictors to assess overall differences in mortality rates
```{r}
# Examining multiple predictors
glm_model <- glm(mortalities ~ Year + race_ethnicity + sex, data = county_race_sex, family = gaussian())
glm_results <- broom::tidy(glm_model)

# Display and save the GLM results
print(glm_results)
saveRDS(glm_results, file = here::here("products", "manuscript", "supplement", "glm_results.rds"))

# Plot GLM predictions
  county_race_sex$predicted_mortalities <- predict(glm_model, type = "response")
  glm_race_sex_plot <- ggplot(county_race_sex, aes(x = Year, y = predicted_mortalities, color = race_ethnicity)) +
    geom_line(size = 1) +
    theme_minimal() +
    labs(title = "GLM Predictions: Mortality Rates by Year and Race/Ethnicity", x = "Year", y = "Predicted Mortality Rate", color = "Race/Ethnicity") +
    scale_x_continuous(breaks = seq(min(county_race_sex$Year, na.rm = TRUE), max(county_race_sex$Year, na.rm = TRUE), by = 5)) +
    theme(legend.position = "right", legend.text = element_text(size = 10), legend.key.size = unit(0.4, "cm"), legend.spacing.y = unit(0.2, "cm"), axis.text.x = element_text(angle = 45, hjust = 1))

  print(glm_race_sex_plot)
  ggsave(here::here("products", "manuscript", "supplement", "glm_race_sex_plot.png"), plot = glm_race_sex_plot)

```

The GLM model incorporating race/ethnicity and sex reveals substantial disparities in mortality rates between groups. Black populations consistently show the highest predicted rates across the period.

####1.3 GLM: Year by Race interaction model:
Fit GLM including interaction between year and race to examine race-specific trends over time

```{r}

#explore whether interactions exist between year and race/sex:

glm_model_interaction <- glm(mortalities ~ Year * race_ethnicity + sex, data = county_race_sex, family = gaussian())

glm_results_interaction <- broom::tidy(glm_model_interaction)
print(glm_results_interaction)

saveRDS(glm_results_interaction, file = here::here("products", "manuscript", "supplement", "glm_results_interaction.rds"))

# Plot the interaction model results
county_race_sex$predicted_mortalities_interaction <- predict(glm_model_interaction, type = "response")
glm_interaction_plot <- ggplot(county_race_sex, aes(x = Year, y = predicted_mortalities_interaction, color = race_ethnicity)) +
  geom_line(size = 1) +
  theme_minimal() +
  labs(title = "GLM Interaction Model: Mortality Trends by Year and Race", x = "Year", y = "Predicted Mortality Rate", color = "Race/Ethnicity") +
  scale_x_continuous(breaks = seq(min(county_race_sex$Year, na.rm = TRUE), max(county_race_sex$Year, na.rm = TRUE), by = 5)) +
  theme(legend.position = "right", legend.text = element_text(size = 10), legend.key.size = unit(0.4, "cm"), legend.spacing.y = unit(0.2, "cm"), axis.text.x = element_text(angle = 45, hjust = 1))

print(glm_interaction_plot)
ggsave(here::here("products", "manuscript", "supplement", "glm_interaction_plot.png"), plot = glm_interaction_plot)

saveRDS(county_race_sex, file = here::here("results", "output", "county_race_sex_with_preds.rds"))
```
By including interaction terms, we can evaluate whether different racial/ethnic groups experienced distinct temporal trends. Here, trends diverge sharply: Black mortality rates decline, while rates rise for American Indian/Alaska Native populations.

####1.4: GLM: Year by Age group

GLM with interaction between year and age group to examine differential temporal trends
```{r}
# GLM with interaction: does time trend differ by age group?
glm_age_group <- glm(mortalities ~ Year * age_group, 
                     data = processed_data_no_percent,
                     family = gaussian())

# Tidy results for review
glm_age_group_results <- broom::tidy(glm_age_group)
print(glm_age_group_results)

# Save results
saveRDS(glm_age_group_results, file = here::here("products", "manuscript", "supplement", "glm_age_group_results.rds"))

# Add predicted values
processed_data_no_percent$predicted_glm_age <- predict(glm_age_group, type = "response")

# Plot actual vs predicted for each age group
glm_age_plot <- ggplot(processed_data_no_percent, 
                       aes(x = Year, y = mortalities, color = age_group)) +
  geom_point(alpha = 0.4) +
  geom_line(aes(y = predicted_glm_age), size = 1.2) +
  theme_minimal() +
  labs(title = "Predicted Mortality Trends by Age Group",
       subtitle = "From GLM: mortalities ~ Year * age_group",
       x = "Year", y = "Mortality Rate") +
  scale_color_brewer(palette = "Dark2")

print(glm_age_plot)

# Save plot
ggsave(here::here("products", "manuscript", "supplement", "glm_age_plot.png"), 
       plot = glm_age_plot, width = 9, height = 6)

```
This model identifies age-specific trends. Mortality among individuals aged 65+ remains substantially higher and shows modest increases, while younger age groups have lower, flatter trends.

####1.5: Slope comparison by Age group

Comparing slopes using the emmeans function.
Use estimated marginal trends (emtrends) to compare yearly slope by age group.

```{r}
# Get slope (trend) estimates for Year, stratified by age group
em_trends_age <- emtrends(glm_age_group, ~ age_group, var = "Year")

# View summary
summary(em_trends_age)
em_trends_age

# compare trends between groups (pairwise contrast)
slope_contrast <- contrast(em_trends_age, method = "pairwise")
summary(slope_contrast)

# Save
saveRDS(em_trends_age, file = here::here("products", "manuscript", "supplement", "emtrends_age_group.rds"))
saveRDS(slope_contrast, file = here::here("products", "manuscript", "supplement", "slope_contrast_age_group.rds"))

# Convert emtrends object to data frame
em_trend_df <- as.data.frame(summary(em_trends_age))

# Create horizontal bar plot with fixed x-axis breaks
slope_plot_agegroup <- ggplot(em_trend_df, aes(x = Year.trend, y = age_group)) +
  geom_col(fill = "#2C77B0", width = 0.5) +
  geom_errorbarh(aes(xmin = Year.trend - 1.96 * SE,
                     xmax = Year.trend + 1.96 * SE),
                 height = 0.15, color = "gray40") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  scale_x_continuous(
    breaks = seq(0, 1.5, by = 0.25),
    limits = c(0, 1.5)
  ) +
  labs(
    title = "Estimated Annual Mortality Rate Change by Age Group",
    subtitle = "Point estimates and 95% confidence intervals",
    x = "Estimated Slope (Yearly Change)",
    y = "Age Group"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 14, margin = margin(b = 6)),
    plot.subtitle = element_text(size = 11, margin = margin(b = 10)),
    axis.text = element_text(size = 12),
    panel.grid.major.y = element_blank())

# Display plot
slope_plot_agegroup

# Save figure to manuscript folder
ggsave(
  filename = here::here("results", "figures", "slope_plot_age_group.png"),
  plot = slope_plot_agegroup,
  width = 9, height = 6, dpi = 300)

# Save RDS for reproducibility
saveRDS(slope_plot_agegroup, file = here::here("results", "figures", "slope_plot_age_group.rds"))

```

Using estimated marginal trends, we formally test whether the yearly rate of change differs between age groups. Older adults (65+) show a significantly steeper upward trend compared to younger adults.

## Aim 2: Spatial Analysis of Hypertension Mortality Rates
## Evaluating spatial distribution and clustering patterns

Evaluating spatial patterns for mortality data for all races together and individually.

###2.1: Static maps by Race:

Converting county data to spatial object using longitude/latitude.
Creating static spatial map for all race groups.
```{r}
#Convert dataset to an SF spatial object
county_race_sex_sf <- county_race_sex %>%
  st_as_sf(coords = c("X_long", "Y_lat"), crs = 4326, remove = FALSE)

# Spatial plot for all races
spatial_plot_all <- ggplot() +
  geom_sf(data = county_race_sex_sf, aes(color = mortalities), size = 0.7) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Spatial Distribution of Mortality Rates (All Races)", x = "Longitude", y = "Latitude", color = "Mortality Rate")

print(spatial_plot_all)
ggsave(here::here("products", "manuscript", "supplement", "spatial_plot_all.png"), plot = spatial_plot_all)

# Spatial plot for Black race
black_race_sf <- county_race_sex_sf %>% filter(race_ethnicity == "Black")
black_spatial_plot <- ggplot() +
  geom_sf(data = black_race_sf, aes(color = mortalities), size = 0.7) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Spatial Distribution of Mortality Rates (Black Race)", x = "Longitude", y = "Latitude", color = "Mortality Rate")

print(black_spatial_plot)
ggsave(here::here("products", "manuscript", "supplement", "black_spatial_plot.png"), plot = black_spatial_plot)

# Spatial plot for White race
white_race_sf <- county_race_sex_sf %>% filter(race_ethnicity == "White")
white_spatial_plot <- ggplot() +
  geom_sf(data = white_race_sf, aes(color = mortalities), size = 0.7) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Spatial Distribution of Mortality Rates (White Race)", x = "Longitude", y = "Latitude", color = "Mortality Rate")

print(white_spatial_plot)
ggsave(here::here("products", "manuscript", "supplement", "white_spatial_plot.png"), plot = white_spatial_plot)

# Spatial plot for Hispanic race
hispanic_sf <- county_race_sex_sf %>% filter(race_ethnicity == "Hispanic")
hispanic_spatial_plot <- ggplot() +
  geom_sf(data = hispanic_sf, aes(color = mortalities), size = 0.7) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Spatial Distribution of Mortality Rates (Hispanic)", x = "Longitude", y = "Latitude", color = "Mortality Rate")

print(hispanic_spatial_plot)
ggsave(here::here("products", "manuscript", "supplement", "hispanic_spatial_plot.png"), plot = hispanic_spatial_plot)

# Spatial plot for American Indian and Alaska Native race
native_sf <- county_race_sex_sf %>% filter(race_ethnicity == "American Indian and Alaska Native")
native_spatial_plot <- ggplot() +
  geom_sf(data = native_sf, aes(color = mortalities), size = 0.7) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Spatial Distribution of Mortality Rates\n(American Indian and Alaska Native)", x = "Longitude", y = "Latitude", color = "Mortality Rate")

print(native_spatial_plot)
ggsave(here::here("products", "manuscript", "supplement", "native_spatial_plot.png"), plot = native_spatial_plot)

# Spatial plot for Asian and Pacific Islander race
asian_sf <- county_race_sex_sf %>% filter(race_ethnicity == "Asian and Pacific Islander")
asian_spatial_plot <- ggplot() +
  geom_sf(data = asian_sf, aes(color = mortalities), size = 0.7) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Spatial Distribution of Mortality Rates\n(Asian and Pacific Islander)", x = "Longitude", y = "Latitude", color = "Mortality Rate")

print(asian_spatial_plot)
ggsave(here::here("products", "manuscript", "supplement", "asian_spatial_plot.png"), plot = asian_spatial_plot)

```

###2.2 Combined spatial plot

Combining all spatial race-specific maps into a multi-panel figure.
```{r}

# Combine all of the spatial charts using patchwork
combined_spatial_plot <- (spatial_plot_all | black_spatial_plot | white_spatial_plot) /
  (hispanic_spatial_plot | native_spatial_plot |
  asian_spatial_plot)

# Print and save the combined spatial plot
print(combined_spatial_plot)
ggsave(here::here("products", "manuscript", "supplement", "combined_spatial_plot.png"), 
       plot = combined_spatial_plot, width = 14, height = 10)

```

Evaluating spatial clustering of Mortality Rates

### 2.3: Global Moran’s I – All Races Combined

We will run a Global Morans analysis to evalaute spatial clustering using K-nearest neighbour clustering. We are looking at the full dataset here which is very computationally intensive. Thus we are setting the code to load the output, but if its not present (not saved) then it will run.
We have saved the output in the large

Running Global Moran's I test to evaluate overall spatial autocorrelation.
```{r spatial-weights global-morans-i, message=FALSE}
# Define file paths for output objrects
knn_path <- here::here("results", "output", "knn_neighbors.rds") # Larger, reusable object
lw_path <- here::here("results", "output", "spatial_weights_list.rds") # Larger, reusable object
moran_path <- here::here("results", "output", "global_morans_i.rds") # Smaller, result-level output

#Reuse or compute spatial weights
#computing is very computationally intensive because we are usign the full dataset

#Check if neighbor and weight objects are already saved
if (file.exists(knn_path) && file.exists(lw_path)) {
   # Load saved spatial weights
  message("Reusing saved spatial weights and neighbors.")
  knn_nb <- readRDS(knn_path)
  lw <- readRDS(lw_path)
} else {
  message("Generating spatial weights from coordinates.")
coords <- st_coordinates(county_race_sex_sf)
knn_nb <- knn2nb(knearneigh(coords, k = 5)) # Create neighbors list - 5 nearest neighbours
lw <- nb2listw(knn_nb, style = "W") # Convert to spatial weights list

  # Save for reuse
  saveRDS(knn_nb, file = knn_path)
  saveRDS(lw, file = lw_path)}

# Global Moran’s I analysis
moran_result <- moran.test(county_race_sex_sf$mortalities, lw)
print(moran_result)

# Save result
saveRDS(moran_result, file = moran_path)

```

Global Moran's I: I = 0.0435, p < 0.001, this indicates weak but statistically significant spatial autocorrelation.
There is weak-to-moderate positive spatial autocorrelation in mortality rates across U.S. counties — counties with high (or low) mortality tend to be near others with similarly high (or low) values. Although the magnitude is small (I = 0.0435), the pattern is considered highly statistically significant based on the p-value.

###2.4: Local Moran’s I (LISA)

Running Local Moran's I (LISA) to detect local spatial clusters.
```{r lisa-analysis, message=FALSE}

# Run Local Moran’s I (LISA)
local_moran <- localmoran(county_race_sex_sf$mortalities, lw)

# Build a clean spatial object with LISA values
county_lisa <- county_race_sex_sf %>%
  st_drop_geometry() %>%
  bind_cols(geometry = st_geometry(county_race_sex_sf)) %>%
  mutate(local_I = local_moran[, 1],     # Local I statistic
         local_p = local_moran[, 5]) %>% # p-value
  st_as_sf()

# Save outputs
saveRDS(local_moran, file = here::here("results", "output", "local_morans_matrix.rds"))
saveRDS(county_lisa, file = here::here("results", "output", "county_race_sex_sf_with_lisa.rds"))
saveRDS(moran_result, here::here("results", "output", "us_global_morans_i.rds"))

```

Creating map of statistically significant LISA clusters (p < 0.05).
```{r}

# Create a data frame with centroids for plotting and significance flag
county_lisa_coords <- county_lisa %>%
  mutate(long = st_coordinates(.)[, 1],
    lat = st_coordinates(.)[, 2],
    significant_cluster = ifelse(local_p < 0.05, "p < 0.05", "Not Significant")) %>%
  st_drop_geometry()

# Plot using points instead of polygons
lisa_point_plot <- ggplot(county_lisa_coords) +
  geom_point(aes(x = long, y = lat, color = significant_cluster), alpha = 0.6, size = 0.7) +
  scale_color_manual(values = c("gray80", "red"),
                     labels = c("Not Significant", "p < 0.05")) +
  labs(title = "Counties with Statistically Significant Clustering",
       color = "LISA Significance") +
  coord_equal() +
  theme_minimal(base_size = 12)

# Save plot
ggsave(filename = here::here("products", "manuscript", "supplement", "lisa_significance_map_centroids.png"),
  plot = lisa_point_plot,
  width = 10,
  height = 6)

print(lisa_point_plot)

```
Classifying counties based on LISA cluster type using z-scores.

```{r classify-lisa-clusters, message=FALSE}
# Compute spatial lag of mortality
spatial_lag <- lag.listw(lw, county_lisa$mortalities)

# Standardize original values and spatial lags (z-scores)
mortality_z <- scale(county_lisa$mortalities)[, 1]
lag_z       <- scale(spatial_lag)[, 1]

# Add z-scores and spatial lag to dataset
county_lisa <- county_lisa %>%
  mutate(
    mortality_z = mortality_z,
    lag_z = lag_z,
    lisa_cluster_type = case_when(
      local_p >= 0.05              ~ "Not Significant",
      mortality_z > 0 & lag_z > 0  ~ "High-High",
      mortality_z < 0 & lag_z < 0  ~ "Low-Low",
      mortality_z > 0 & lag_z < 0  ~ "High-Low (Outlier)",
      mortality_z < 0 & lag_z > 0  ~ "Low-High (Outlier)",
      TRUE                         ~ NA_character_))

```

Generating map of LISA cluster types (High-High, Low-Low, Outliers).

```{r lisa-cluster-type-map, message=FALSE}
# Create centroid-based coordinates for plotting
county_lisa_coords <- county_lisa %>%
  mutate(long = st_coordinates(.)[, 1],
         lat = st_coordinates(.)[, 2]) %>%
  st_drop_geometry()

# Define cluster color palette
cluster_colors <- c(
  "High-High" = "red3",
  "Low-Low" = "blue3",
  "High-Low (Outlier)" = "orange",
  "Low-High (Outlier)" = "purple",
  "Not Significant" = "gray85"
)

# Plot cluster types
lisa_cluster_plot <- ggplot(county_lisa_coords) +
  geom_point(aes(x = long, y = lat, color = lisa_cluster_type), alpha = 0.7, size = 0.75) +
  scale_color_manual(values = cluster_colors) +
  labs(
    title = "Local Spatial Clusters of Mortality (LISA)",
    subtitle = "High–High = Hotspot, Low–Low = Coldspot, Others = Outliers",
    color = "Cluster Type"
  ) +
  coord_equal() +
  theme_minimal(base_size = 12)

# Print and save
print(lisa_cluster_plot)

ggsave(
  here::here("products", "manuscript", "supplement", "lisa_cluster_type_map.png"),
  plot = lisa_cluster_plot,
  width = 10, height = 6
)

```
Local Moran’s I was computed on the full national dataset using 5-nearest-neighbor spatial weights. Each county was represented by a centroid, and clustering patterns reflect spatial autocorrelation of mortality rates across all years and subgroups.

####Focusing at a State level
Due to the size of the dataset it is challenging to work with the full dataset.
For convenience and because the SouthEast is historically known to have a high prevalence of cardiovasccular mortalities we will focus on Georgia for this next step.
Our previous analysis support that there is a general distribution of cases here, even if there are not distinct hotspots.

Analyzing spatial clusters in Georgia using LISA.
```{r}
# Georgia: Create or load spatial weights
ga_knn_path <- here::here("results", "output", "ga_knn_neighbors.rds")
ga_lw_path  <- here::here("results", "output", "ga_spatial_weights.rds")

county_ga_sf <- county_race_sex_sf %>%
  filter(LocationAbbr == "GA")

if (file.exists(ga_knn_path) && file.exists(ga_lw_path)) {
  message("Reusing saved GA spatial weights and neighbors.")
  ga_knn_nb <- readRDS(ga_knn_path)
  ga_lw <- readRDS(ga_lw_path)
} else {
  message("Generating GA spatial weights.")
  ga_coords <- st_coordinates(county_ga_sf)
  ga_knn_nb <- knn2nb(knearneigh(ga_coords, k = 5))
  ga_lw     <- nb2listw(ga_knn_nb, style = "W")
  saveRDS(ga_knn_nb, file = ga_knn_path)
  saveRDS(ga_lw, file = ga_lw_path)
}

```

```{r ga-lisa-analysis, message=FALSE}

# Global Moran's I
ga_moran_result <- moran.test(county_ga_sf$mortalities, ga_lw)
print(ga_moran_result)
saveRDS(ga_moran_result, here::here("results", "output", "ga_global_morans_i.rds"))

# Local Moran's I (LISA)
ga_local_moran <- localmoran(county_ga_sf$mortalities, ga_lw)

# Create cluster type classification
ga_spatial_lag <- lag.listw(ga_lw, county_ga_sf$mortalities)
ga_z <- scale(county_ga_sf$mortalities)[,1]
ga_lag_z <- scale(ga_spatial_lag)[,1]
ga_p <- ga_local_moran[,5]

county_ga_sf <- county_ga_sf %>%
  mutate(
    local_I = ga_local_moran[, 1],
    local_p = ga_p,
    mortality_z = ga_z,
    lag_z = ga_lag_z,
    cluster_type = case_when(
      local_p >= 0.05 ~ "Not Significant",
      mortality_z > 0 & lag_z > 0 ~ "High-High",
      mortality_z < 0 & lag_z < 0 ~ "Low-Low",
      mortality_z > 0 & lag_z < 0 ~ "High-Low (Outlier)",
      mortality_z < 0 & lag_z > 0 ~ "Low-High (Outlier)",
      TRUE ~ NA_character_))

# Plot cluster types using centroids
ga_coords_df <- county_ga_sf %>%
  mutate(long = st_coordinates(.)[,1],
         lat = st_coordinates(.)[,2]) %>%
  st_drop_geometry()

cluster_colors <- c(
  "High-High" = "red3",
  "Low-Low" = "blue3",
  "High-Low (Outlier)" = "orange",
  "Low-High (Outlier)" = "purple",
  "Not Significant" = "gray80"
)

ga_lisa_plot <- ggplot(ga_coords_df) +
  geom_point(aes(x = long, y = lat, color = cluster_type), alpha = 0.8, size = 2) +
  scale_color_manual(values = cluster_colors) +
  labs(
    title = "LISA Cluster Types: Georgia",
    subtitle = "High–High = Hotspot; Low–Low = Coldspot; Others = Outliers",
    color = "Cluster Type"
  ) +
  coord_equal() +
  theme_minimal(base_size = 13)

# Display and save
print(ga_lisa_plot)

ggsave(
  here::here("products", "manuscript", "supplement", "ga_lisa_cluster_plot.png"),
  plot = ga_lisa_plot,
  width = 8,
  height = 6)
```

A spatial weights matrix was created for Georgia counties using nearest-neighbor distances. LISA statistics were computed to explore clustering, but no statistically significant local hotspots were observed.

Analyzing spatial clusters in Mississippi using LISA.

```{r ms-lisa-analysis, message=FALSE}

# Define file paths for MS spatial objects
ms_knn_path <- here::here("results", "output", "ms_knn_neighbors.rds")
ms_lw_path  <- here::here("results", "output", "ms_spatial_weights.rds")
ms_moran_path <- here::here("results", "output", "ms_global_morans_i.rds")

# Subset to Mississippi (MS)
county_ms_sf <- county_race_sex_sf %>%
  filter(LocationAbbr == "MS")

# Reuse or generate new spatial weights
if (file.exists(ms_knn_path) && file.exists(ms_lw_path)) {
  message("Reusing saved MS spatial weights and neighbors.")
  ms_knn_nb <- readRDS(ms_knn_path)
  ms_lw     <- readRDS(ms_lw_path)
} else {
  message("Generating MS spatial weights.")
  ms_coords <- st_coordinates(county_ms_sf)
  ms_knn_nb <- knn2nb(knearneigh(ms_coords, k = 5))
  ms_lw     <- nb2listw(ms_knn_nb, style = "W")

  # Save for reuse
  saveRDS(ms_knn_nb, file = ms_knn_path)
  saveRDS(ms_lw, file = ms_lw_path)
}

# Global Moran’s I for MS
ms_moran_result <- moran.test(county_ms_sf$mortalities, ms_lw)
print(ms_moran_result)
saveRDS(ms_moran_result, file = ms_moran_path)

# Local Moran’s I (LISA)
ms_local_moran <- localmoran(county_ms_sf$mortalities, ms_lw)

# Classify cluster types
ms_spatial_lag <- lag.listw(ms_lw, county_ms_sf$mortalities)
ms_z <- scale(county_ms_sf$mortalities)[,1]
ms_lag_z <- scale(ms_spatial_lag)[,1]
ms_p <- ms_local_moran[,5]

county_ms_sf <- county_ms_sf %>%
  mutate(
    local_I = ms_local_moran[, 1],
    local_p = ms_p,
    mortality_z = ms_z,
    lag_z = ms_lag_z,
    cluster_type = case_when(
      local_p >= 0.05 ~ "Not Significant",
      mortality_z > 0 & lag_z > 0 ~ "High-High",
      mortality_z < 0 & lag_z < 0 ~ "Low-Low",
      mortality_z > 0 & lag_z < 0 ~ "High-Low (Outlier)",
      mortality_z < 0 & lag_z > 0 ~ "Low-High (Outlier)",
      TRUE ~ NA_character_
    )
  )

# Plot with centroids
ms_coords_df <- county_ms_sf %>%
  mutate(long = st_coordinates(.)[,1],
         lat = st_coordinates(.)[,2]) %>%
  st_drop_geometry()

cluster_colors <- c(
  "High-High" = "red3",
  "Low-Low" = "blue3",
  "High-Low (Outlier)" = "orange",
  "Low-High (Outlier)" = "purple",
  "Not Significant" = "gray80"
)

ms_lisa_plot <- ggplot(ms_coords_df) +
  geom_point(aes(x = long, y = lat, color = cluster_type), alpha = 0.8, size = 2.5) +
  scale_color_manual(values = cluster_colors) +
  labs(
    title = "LISA Cluster Types: Mississippi",
    subtitle = "High–High = Hotspot; Low–Low = Coldspot; Others = Outliers",
    color = "Cluster Type"
  ) +
  coord_equal() +
  theme_minimal(base_size = 13)

print(ms_lisa_plot)

ggsave(
  filename = here::here("products", "manuscript", "supplement", "ms_lisa_cluster_plot.png"),
  plot = ms_lisa_plot,
  width = 8,
  height = 6
)

```

LISA analysis for Mississippi counties detected spatial relationships using 5-nearest-neighbor weights. Similar to other states, no prominent local clusters were found despite elevated mortality rates in some areas.

Analyzing spatial clusters in Texas using LISA.

```{r tx-lisa-analysis, message=FALSE}

# Define file paths for Texas-specific spatial objects
tx_knn_path <- here::here("results", "output", "tx_knn_neighbors.rds")
tx_lw_path  <- here::here("results", "output", "tx_spatial_weights.rds")
tx_moran_path <- here::here("results", "output", "tx_global_morans_i.rds")

# Subset to Texas counties
county_tx_sf <- county_race_sex_sf %>%
  filter(LocationAbbr == "TX")

# Reuse or generate neighbors/weights
if (file.exists(tx_knn_path) && file.exists(tx_lw_path)) {
  message("Reusing saved TX spatial weights and neighbors.")
  tx_knn_nb <- readRDS(tx_knn_path)
  tx_lw     <- readRDS(tx_lw_path)
} else {
  message("Generating TX spatial weights.")
  tx_coords <- st_coordinates(county_tx_sf)
  tx_knn_nb <- knn2nb(knearneigh(tx_coords, k = 5))
  tx_lw     <- nb2listw(tx_knn_nb, style = "W")

  # Save for reuse
  saveRDS(tx_knn_nb, file = tx_knn_path)
  saveRDS(tx_lw, file = tx_lw_path)
}

```

```{r tx-lisa-analysis, message=FALSE}

# Global Moran’s I for Texas
tx_moran_result <- moran.test(county_tx_sf$mortalities, tx_lw)
print(tx_moran_result)
saveRDS(tx_moran_result, file = tx_moran_path)

# Local Moran’s I (LISA)
tx_local_moran <- localmoran(county_tx_sf$mortalities, tx_lw)

# Spatial lag and z-scores
tx_spatial_lag <- lag.listw(tx_lw, county_tx_sf$mortalities)
tx_z <- scale(county_tx_sf$mortalities)[,1]
tx_lag_z <- scale(tx_spatial_lag)[,1]
tx_p <- tx_local_moran[,5]

# Add LISA values and classification
county_tx_sf <- county_tx_sf %>%
  mutate(
    local_I = tx_local_moran[, 1],
    local_p = tx_p,
    mortality_z = tx_z,
    lag_z = tx_lag_z,
    cluster_type = case_when(
      local_p >= 0.05 ~ "Not Significant",
      mortality_z > 0 & lag_z > 0 ~ "High-High",
      mortality_z < 0 & lag_z < 0 ~ "Low-Low",
      mortality_z > 0 & lag_z < 0 ~ "High-Low (Outlier)",
      mortality_z < 0 & lag_z > 0 ~ "Low-High (Outlier)",
      TRUE ~ NA_character_
    )
  )

# Convert to points (centroids) for visualization
tx_coords_df <- county_tx_sf %>%
  mutate(long = st_coordinates(.)[,1],
         lat = st_coordinates(.)[,2]) %>%
  st_drop_geometry()

# Cluster colors
cluster_colors <- c(
  "High-High" = "red3",
  "Low-Low" = "blue3",
  "High-Low (Outlier)" = "orange",
  "Low-High (Outlier)" = "purple",
  "Not Significant" = "gray80"
)

# Plot
tx_lisa_plot <- ggplot(tx_coords_df) +
  geom_point(aes(x = long, y = lat, color = cluster_type), alpha = 0.8, size = 2) +
  scale_color_manual(values = cluster_colors) +
  labs(
    title = "LISA Cluster Types: Texas",
    subtitle = "High–High = Hotspot; Low–Low = Coldspot; Others = Outliers",
    color = "Cluster Type"
  ) +
  coord_equal() +
  theme_minimal(base_size = 13)

print(tx_lisa_plot)

# Save plot
ggsave(
  filename = here::here("products", "manuscript", "supplement", "tx_lisa_cluster_plot.png"),
  plot = tx_lisa_plot,
  width = 9, height = 6
)
```

LISA analysis was performed on counties in Texas using centroid-based spatial weights. While global spatial autocorrelation was detected, no statistically significant local clusters were identified.

Analyzing spatial clusters in Kentucky using LISA.

```{r}

# Paths for saving spatial weights
ky_knn_path <- here::here("results", "output", "ky_knn_neighbors.rds")
ky_lw_path  <- here::here("results", "output", "ky_spatial_weights.rds")

# Subset to Kentucky counties
county_ky_sf <- county_race_sex_sf %>%
  filter(LocationAbbr == "KY")

# Reuse or generate spatial weights for Kentucky
if (file.exists(ky_knn_path) && file.exists(ky_lw_path)) {
  message("Reusing saved KY spatial weights and neighbors.")
  ky_knn_nb <- readRDS(ky_knn_path)
  ky_lw     <- readRDS(ky_lw_path)
} else {
  message("Generating KY spatial weights.")
  ky_coords <- st_coordinates(county_ky_sf)
  ky_knn_nb <- knn2nb(knearneigh(ky_coords, k = 5))
  ky_lw     <- nb2listw(ky_knn_nb, style = "W")

  # Save for reuse
  saveRDS(ky_knn_nb, file = ky_knn_path)
  saveRDS(ky_lw, file = ky_lw_path)
}

# Compute local Moran’s I
local_moran_ky <- localmoran(county_ky_sf$mortalities, ky_lw)

# Add results to object
county_ky_sf <- county_ky_sf %>%
  mutate(
    local_I = local_moran_ky[, 1],
    local_p = local_moran_ky[, 5],
    mortality_z = scale(mortalities)[, 1],
    lag_z = lag.listw(ky_lw, mortality_z),
    lisa_cluster = case_when(
      local_p > 0.05 ~ "Not Significant",
      mortality_z > 0 & lag_z > 0 ~ "High-High",
      mortality_z < 0 & lag_z < 0 ~ "Low-Low",
      mortality_z > 0 & lag_z < 0 ~ "High-Low (Outlier)",
      mortality_z < 0 & lag_z > 0 ~ "Low-High (Outlier)",
      TRUE ~ NA_character_
    )
  )

# Get centroid coordinates
county_ky_coords <- county_ky_sf %>%
  mutate(long = st_coordinates(.)[, 1],
         lat = st_coordinates(.)[, 2]) %>%
  st_drop_geometry()

# Plot
lisa_ky_plot <- ggplot(county_ky_coords) +
  geom_point(aes(x = long, y = lat, color = lisa_cluster), alpha = 0.7, size = 0.8) +
  scale_color_manual(
    values = c("High-High" = "red", "Low-Low" = "blue",
               "High-Low (Outlier)" = "orange", "Low-High (Outlier)" = "purple", "Not Significant" = "gray80"),
    na.value = "black"
  ) +
  labs(
    title = "LISA Cluster Types: Kentucky",
    subtitle = "High–High = Hotspot; Low–Low = Coldspot; Others = Outliers",
    color = "Cluster Type"
  ) +
  coord_equal() +
  theme_minimal(base_size = 12)

# Save
ggsave(
  here::here("products", "manuscript", "supplement", "lisa_cluster_kentucky.png"),
  plot = lisa_ky_plot, width = 9, height = 6)

# Compute Global Moran’s I for Kentucky
ky_moran_result <- moran.test(county_ky_sf$mortalities, ky_lw)

# Save the result
saveRDS(ky_moran_result, here::here("results", "output", "ky_global_morans_i.rds"))

print(lisa_ky_plot)

```
the LISA plot shows spatial patterns across all available data, not just one year

Compile Global Moran’s I test results from the USA and selected high-mortality states (GA, MS, TX, KY).
This table summarizes the spatial autocorrelation of mortality rates.

2.5 Global Moran's I Summary Table

```{r}
# Already saved or placed earlier i
extract_moran_values <- function(obj, geo) {
  tibble(
    Geography = geo,
    `Moran's I` = round(obj$estimate[["Moran I statistic"]], 4),
    Expectation = round(obj$estimate[["Expectation"]], 4),
    Variance = round(obj$estimate[["Variance"]], 6),
    `Z Statistic` = round(obj$statistic, 4),
    `p-value` = signif(obj$p.value, 4))}

# Load all Moran's I results
usa_moran <- readRDS(here::here("results", "output", "global_morans_i.rds"))
ga_moran  <- readRDS(here::here("results", "output", "ga_global_morans_i.rds"))
ms_moran  <- readRDS(here::here("results", "output", "ms_global_morans_i.rds"))
tx_moran  <- readRDS(here::here("results", "output", "tx_global_morans_i.rds"))
ky_moran  <- readRDS(here::here("results", "output", "ky_global_morans_i.rds"))

# Build summary table
moran_summary <- bind_rows(
  extract_moran_values(usa_moran, "USA (All Counties)"),
  extract_moran_values(ga_moran,  "Georgia"),
  extract_moran_values(ms_moran,  "Mississippi"),
  extract_moran_values(tx_moran,  "Texas"),
  extract_moran_values(ky_moran,  "Kentucky"))

# Format with gt
moran_gt_tbl <- moran_summary %>%
  gt() %>%
  tab_header(
    title = "Global Moran's I Statistics by Geography",
    subtitle = "Spatial Autocorrelation of Mortality Rates"
  ) %>%
  fmt_number(columns = 2:5, decimals = 4) %>%  # Keep decimals for other metrics
  fmt_scientific(columns = vars(`p-value`), decimals = 2) %>%  # Use sci notation for p-values
  cols_label(
    `Moran's I` = "Moran's I",
    `Z Statistic` = "Z Statistic",
    `p-value` = "p-value"
  ) %>%
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12),
    column_labels.font.weight = "bold"
  )

# Save table
saveRDS(moran_gt_tbl, here::here("results", "tables", "moran_global_summary.rds"))

# View
moran_gt_tbl
```

A summary table of Global Moran’s I statistics for the full U.S. dataset and selected states (GA, MS, TX, KY) is shown below, indicating statistically significant spatial autocorrelation in all geographies examined (p < 0.001).

Local Indicators of Spatial Association (LISA) – Statistically Significant Clustering of Mortality Rates in Kentucky.
This map highlights counties with statistically significant spatial clustering of hypertension-related cardiovascular disease (CVD) mortality rates, as detected by Local Moran’s I. Each point represents a county centroid. Counties shown in red are spatial clusters where both the county and its neighbors have high mortality values (high-high) or display significant spatial autocorrelation (e.g., high-low or low-high). Spatial relationships were calculated using 5-nearest-neighbor spatial weights based on county centroids. This analysis includes data from all available years and demographic subgroups; therefore, significance reflects elevated spatial correlation within any combination of groupings, not just overall rates.

Local spatial analysis (LISA) identified a small number of statistically significant clusters in Kentucky, including high-mortality counties adjacent to other high-mortality areas. These patterns support the presence of localized mortality hotspots and may inform targeted investigation or intervention efforts.

##Aim 3: Demographic Disparities

### 3.1: Estimated sloped by Race and Sex

Slope Comparison by Race and Sex using emmeans function for all ages.

Extract estimated annual mortality trends by race/ethnicity and sex across all adults (35+). These values represent the modeled rate of change in mortality per year.

```{r}
# Extract estimated yearly trends (slopes) by race_ethnicity and sex
em_trends_race_sex <- emtrends(glm_model_interaction, ~ race_ethnicity + sex, var = "Year")

# Summarize estimated slopes
# View slope estimates (with SE and CI) for each group
summary(em_trends_race_sex)
print(em_trends_race_sex)

# Compare slopes between race/sex groups
slope_contrast_race_sex <- contrast(em_trends_race_sex, method = "pairwise")

# View significance of differences in slopes
summary(slope_contrast_race_sex)
print(slope_contrast_race_sex)

# Save slope estimates and pairwise contrasts
saveRDS(em_trends_race_sex, file = here::here("results", "output", "emtrends_race_sex.rds"))
saveRDS(slope_contrast_race_sex, file = here::here("results", "output", "slope_contrast_race_sex.rds"))

# Plot slope estimates
slope_plot_race_sex <- plot(em_trends_race_sex) +
  ggtitle("Estimated Annual Mortality Change by Race and Sex") +
  ylab("Yearly Slope Estimate") +
  xlab("Race and Sex Group") +
  theme_minimal(base_size = 12)

# Save and display the plot
ggsave(here::here("results", "output", "slope_plot_race_sex.png"),
       plot = slope_plot_race_sex, width = 10, height = 6)
print(slope_plot_race_sex)
```

Estimated yearly mortality rate trends were computed for each race and sex group. Positive slope estimates indicate increasing mortality over time, while negative values suggest improvements. Pairwise comparisons were used to assess whether these differences were statistically significant across groups. This was for all ages aged 35+ (both groups combined).

###3.2 Demogrpahic disparities by Age group

We now wish to use a model to extract yearly mortality trends by combinations of race/ethnicity and sex while adjusting for age group to see the impact of demosgraphic disparities between age groups. To assess disparities across age, we fit a GLM including three-way interactions between year, race/ethnicity, and sex, adjusting for age group. This allows age-stratified analysis of temporal trends.

Fit the Interaction Model
```{r}
# GLM with 3-way interaction to explore mortality trends by Year * race_ethnicity * sex, stratified by age group

## ---- load_glm_race_sex_age_model, message=TRUE, warning=FALSE -------------------

glm_path <- here::here("results", "output", "glm_race_sex_age_model.rds")

if (file.exists(glm_path)) {
  message("Loading saved GLM model with 3-way interaction...")
  glm_race_sex_age <- readRDS(glm_path)
} else {
  message("Fitting GLM model with 3-way interaction...")
  glm_race_sex_age <- glm(mortalities ~ Year * race_ethnicity * sex + age_group,
                          data = county_race_sex, family = gaussian())
  saveRDS(glm_race_sex_age, glm_path)
}

#glm_race_sex_age <- glm(mortalities ~ Year * race_ethnicity * sex + age_group, data = processed_data_no_percent, family = gaussian())

# Save model
#saveRDS(glm_race_sex_age, here::here("results", "output", "glm_race_sex_age_model.rds"))

```

We fit a GLM with 3-way interactions between year, race/ethnicity, and sex, while adjusting for age group. This allows us to examine mortality trends by race and sex, separately for younger and older adults.

Get Estimated Slopes Stratified by Age Group
```{r}
# Extract yearly slopes within each race-sex group, stratified by age_group
em_trends_race_sex_age <- emtrends(glm_race_sex_age,
                                   ~ race_ethnicity + sex | age_group,
                                   var = "Year")

# Save slopes
saveRDS(em_trends_race_sex_age, here::here("results", "output", "emtrends_race_sex_by_age.rds"))

# Print to console
summary(em_trends_race_sex)
```

This extracts slope estimates (rate of annual change in mortality) for each demographic subgroup, separately for younger and older adults. Estimated yearly changes in mortality rates were extracted for each race/sex group, stratified by age. This approach highlights disparities not only across race and sex but also how these trends differ across age strata.

###3.3 Pairwise Comparisons Within Each Age Group

Compare estimated slopes between all demographic subgroups within each age group using pairwise contrasts.

```{r}
# Compare slopes between race/sex groups within each age group
slope_contrast_race_sex_age <- contrast(em_trends_race_sex_age, method = "pairwise")

# Save
saveRDS(slope_contrast_race_sex_age, here::here("results", "output", "slope_contrast_race_sex_by_age.rds"))

```

Pairwise comparisons were conducted to formally test differences in temporal mortality trends between groups within each age group. These contrasts help assess whether slope differences between groups are statistically significant within each age group.

###3.4: Visualize Slope Contrasts
Top slope differences (absolute magnitude) were visualized to highlight which subgroup comparisons showed the greatest disparities. Significant differences (p < 0.05) are shown in red.
```{r}
# Summarize the contrast object
contrast_summary <- as.data.frame(summary(slope_contrast_race_sex_age)) %>%
  filter(!is.na(estimate)) %>%  # remove any NAs if they exist
  mutate(
    significant = p.value < 0.05,
    contrast = factor(contrast))

contrast_top <- contrast_summary %>%
  arrange(desc(abs(estimate))) %>%
  group_by(age_group) %>%
  slice_head(n = 20) %>%   # Top 20 slope differences per age group
  ungroup()

# Plot
contrast_plot <- ggplot(contrast_top, aes(x = estimate, y = reorder(contrast, estimate), color = significant)) +
  geom_point(size = 2) +
  geom_errorbarh(aes(xmin = estimate - 1.96 * SE, xmax = estimate + 1.96 * SE), height = 0.3) +
  facet_wrap(~ age_group, scales = "free_y") +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "gray40")) +
  labs(
    title = "Top Pairwise Differences in Annual Mortality Trends by Race and Sex",
    x = "Estimated Difference in Yearly Mortality Slope",
    y = "Group Comparison",
    color = "Significant (p < 0.05)") +
  theme_minimal(base_size = 11)

# Save and display
ggsave(
  here::here("results", "output", "pairwise_slope_contrasts_filtered.png"),
  plot = contrast_plot, width = 11, height = 8)
print(contrast_plot)

```

To improve clarity, cleaned labels were applied to simplify group names and facet plots were split by age group. Bar height reflects annual change in mortality; error bars represent 95% confidence intervals.

```{r}
# Convert estimated emtrends object to a data frame for plotting
# Create a group label that includes race, sex, and age group
slope_df <- as.data.frame(em_trends_race_sex_age) %>%
  mutate(group = paste(race_ethnicity, sex),
         group = paste(group, age_group, sep = "___"))

# Helper function to clean group labels (removes age group text)
clean_labels <- function(df) {
  df %>%
    mutate(
      group = str_remove(group, "___Ages 35-64 years"),
      group = str_remove(group, "___Ages 65\\+ years"),
      group = str_replace(group, "Asian and Pacific Islander", "Asian & Pacific\nIslander"),
      group = str_replace(group, "American Indian and Alaska Native", "American Indian\n& Alaska Native"),
      group = str_replace(group, "Overall Female", "Female"),
      group = str_replace(group, "Overall Male", "Male"),
      group = str_replace(group, "Overall Overall", "All Groups"),
      group = str_replace(group, "White Overall", "White"),
      group = str_replace(group, "Black Overall", "Black"),
      group = str_replace(group, "Hispanic Overall", "Hispanic"),
      group = str_replace(group, "Asian & Pacific\nIslander Overall", "Asian & Pacific\nIslander"),
      group = str_replace(group, "American Indian\n& Alaska Native Overall", "American Indian\n& Alaska Native"),
      group = str_trim(group))}

# Apply and split
slope_df_35_64 <- slope_df %>% filter(age_group == "Ages 35-64 years") %>% clean_labels()
slope_df_65_plus <- slope_df %>% filter(age_group == "Ages 65+ years") %>% clean_labels()

# Common ggplot function
make_slope_plot <- function(data, title, fill_color) {
  data <- data %>% 
    filter(!is.na(group), !is.na(Year.trend)) %>%   # remove rows with missing group or slope
    mutate(group = as.factor(group))                # make sure group is a factor

  ggplot(data, aes(x = Year.trend, y = fct_reorder(group, Year.trend))) +
    geom_col(fill = fill_color) +
    geom_errorbarh(aes(xmin = Year.trend - 1.96 * SE,
                       xmax = Year.trend + 1.96 * SE),
                   height = 0.3, color = "black", alpha = 0.7) +
    labs(
      title = title,
      x = "Estimated Annual Change in Mortality Rate",
      y = "Race and Ethnicity"
    ) +
    theme_minimal(base_size = 13) +
    theme(
      plot.title = element_text(hjust = 0.5),
      axis.text.y = element_text(size = 11),
      axis.text.x = element_text(size = 11))}

# Generate plots
slope_plot_35_64 <- make_slope_plot(slope_df_35_64, "Estimated Annual Mortality Change (Ages 35–64)", "#F8766D")
slope_plot_65_plus <- make_slope_plot(slope_df_65_plus, "Estimated Annual Mortality Change (Ages 65+)", "#00BFC4")

# Save plots
ggsave(here("results", "figures", "slope_plot_35_64.png"),
       slope_plot_35_64, width = 11, height = 7, dpi = 300)

ggsave(here("results", "figures", "slope_plot_65_plus.png"),
       slope_plot_65_plus, width = 11, height = 7, dpi = 300)

# Save RDS objects (optional)
saveRDS(slope_plot_35_64, here("results", "figures", "slope_plot_35_64.rds"))
saveRDS(slope_plot_65_plus, here("results", "figures", "slope_plot_65_plus.rds"))

slope_plot_35_64
slope_plot_65_plus
```

Estimated annual change in hypertension-related CVD mortality per 100,000 adults, stratified by race/ethnicity and age group (2000–2019). Bars represent modeled slope estimates, and error bars represent 95% confidence intervals. “Overall” groups include both sexes; sex-specific estimates were excluded where modeling was not feasible.

We assessed yearly changes in hypertension-related cardiovascular disease (CVD) mortality across race/ethnicity and sex groups, stratified by age (35–64 and 65+ years). Modeled slope estimates revealed significant disparities in temporal mortality trends. Black individuals consistently showed the largest reductions in mortality over time, while males and American Indian/Alaska Native populations experienced the steepest increases. These patterns were consistent across both age groups. Pairwise comparisons highlighted statistically significant differences between subgroups, especially between Black adults and other demographic groups. Overall, these findings indicate that improvements in CVD mortality have not been shared equally across populations and age groups.

##Aim 4: Predict County-level Mortality rates

We will make use of some models to try to predict mortality rates based on the different potential predictors (location, race, sex, age-group and year)

####Data Preparation for modeling

```{r}
# Set seed for reproducibility
set.seed(123)

# Use the cleaned modeling dataset
model_data <- county_race_sex %>%
  drop_na(mortalities, Year, age_group, race_ethnicity, sex)  # Remove rows with missing key variables

# Initial train/test split (80/20) - test split stratified by outcome
data_split <- initial_split(model_data, prop = 0.8, strata = mortalities)
train_data <- training(data_split)
test_data  <- testing(data_split)

# 10-fold cross-validation folds (stratified by outcome)
cv_folds <- vfold_cv(train_data, v = 10, strata = mortalities)

```

Generating the processing recipe and model workflow

```{r}
# Preprocessing recipe
mortality_recipe <- recipe(mortalities ~ Year + age_group + race_ethnicity + sex, data = train_data) %>%
  step_dummy(all_nominal_predictors()) %>%  # Convert categorical predictors to dummies
  step_zv(all_predictors()) %>%             # Remove zero-variance predictors (if any)
  step_normalize(all_numeric_predictors())  # Normalize numeric predictors (being cautious)

# Define random forest model
rf_model <- rand_forest(
  mtry = tune(),     # number of variables to try at each split 
  trees = 500,       # total number of trees
  min_n = tune()     # minimum data points in a node 
) %>%
  set_engine("ranger", importance = "impurity") %>%  # impurity-based variable importance
  set_mode("regression")

# Combine into a workflow
rf_workflow <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(mortality_recipe)

```

Hyperparamter tuning and fitting a random forest model:
We will tune a random forest with 10-Fold cross-validation:

# Define path to save tuning results

```{r}
# Define file paths for saving/loading tuning results
tuned_path     <- here::here("results", "large-files", "rf_tuned.rds")
metrics_path   <- here::here("results", "output", "rf_tuned_metrics.rds")
best_param_path <- here::here("results", "output", "rf_best_params.rds")

# Check if tuning results already exist
if (file.exists(tuned_path) && file.exists(metrics_path) && file.exists(best_param_path)) {
  message( "Tuning results already saved. Loading RF tuning results...")

  # Load saved objects
  rf_tuned <- readRDS(tuned_path)
  rf_tuned_metrics <- readRDS(metrics_path)
  best_rf_params <- readRDS(best_param_path)

} else {
  message(" Running Random Forest tuning...")
# Set up a small tuning grid for efficiency
rf_grid <- grid_regular(
  mtry(range = c(1, 4)),     # Explore 1 to 4 predictors
  min_n(range = c(5, 25)),   # Minimum observations per leaf
  levels = 3)                # Total levels in grid for each param

# Tune the random forest model
set.seed(123)
rf_tuned <- tune_grid(
  rf_workflow,
  resamples = cv_folds,
  grid = rf_grid,
  metrics = metric_set(rmse, rsq, mae),
  control = control_grid(save_pred = TRUE))

# Collect metrics from the tuning process
rf_tuned_metrics <- collect_metrics(rf_tuned)
best_rf_params <- select_best(rf_tuned, metric = "rmse")

# Save results to avoid re-running this chunk
  saveRDS(rf_tuned, tuned_path)
  saveRDS(rf_tuned_metrics, metrics_path)
  saveRDS(best_rf_params, best_param_path)

  message("Tuning complete and results saved.")}

```
```{r}
## ---- fit_or_load_rf_model, message=TRUE, warning=FALSE -------------------

use_slim_model <- FALSE # Toggle this to FALSE to regenerate slim model if needed

slim_rf_path <- here::here("results", "output", "rf_model_minimal.rds")
full_rf_path <- here::here("results", "large-files", "rf_tuned.rds")
full_rf_metrics_path <- here::here("results", "output", "rf_tuned_metrics.rds")
best_rf_params_path <- here::here("results", "output", "rf_best_params.rds")

if (use_slim_model) {
  if (!file.exists(slim_rf_path)) {
    stop("Slimmed model not found. Please run with use_slim_model <- FALSE first to generate it.")
  }

  message("Loading slimmed RF model...")
  rf_slim <- readRDS(slim_rf_path)
  rf_final_model <- rf_slim$model
  rf_metrics <- rf_slim$metrics
  rf_vip <- rf_slim$vip
  best_rf_params <- readRDS(best_rf_params_path)

} else {
  if (!file.exists(full_rf_path)) {
    stop("Full Random Forest model not found. Please rerun the tuning block.")
  }

  message("Loading full RF model and generating slim version...")
  rf_tuned <- readRDS(full_rf_path)
  rf_metrics <- readRDS(full_rf_metrics_path)
  rf_final_model <- extract_workflow(rf_tuned)
  rf_vip <- extract_fit_parsnip(rf_final_model) %>% vip::vi()
  best_rf_params <- readRDS(best_rf_params_path)

  if (!file.exists(slim_rf_path)) {
    saveRDS(
      list(model = rf_final_model, metrics = rf_metrics, vip = rf_vip),
      file = slim_rf_path
    )
    message("Slimmed model saved.")}}

```
Creeating a slimmed down version of the tuning RDS file to reduce file size for Github and working with if statements to load this from the correct folder and to not overwrite files.

The slim model can be toggled on or off here by changing from TRUE to FALSE if wanting to use the full version. The full RDS file is saved in the large-files folder which will not automatically upload to github.

```{r}
#Slim version already loads rf_metrics, so no need to reload
rf_metrics %>% filter(.metric == "rmse") %>% arrange(mean)

```

from this we see that the best model based on the lowest RMSE is the first one listed (Preprocessor1_Model6) with all 4 predictors used (mtry = 4), 15 observations per leaf in the tree ( min_n = 15), and RMSE = 115.22.

```{r}

# Plot RMSE across combinations
rf_metrics %>% filter(.metric == "rmse") %>%
  ggplot(aes(x = factor(mtry), y = mean, color = factor(min_n))) +
  geom_point(size = 3) +
  geom_line(aes(group = min_n)) +
  labs(title = "Random Forest Tuning Results (RMSE)",
       x = "mtry (Number of Predictors)",
       y = "Mean RMSE",
       color = "min_n") +
  theme_minimal()

```

```{r}
# File paths for saving
fit_path       <- here::here("results", "output", "final_rf_fit.rds")
workflow_path  <- here::here("results", "output", "final_rf_workflow.rds")
pred_path      <- here::here("results", "output", "rf_test_predictions.rds")
metrics_path   <- here::here("results", "output", "rf_test_metrics.rds")

# Skip if already saved
if (file.exists(fit_path) && file.exists(pred_path)) {
  message("Final model and predictions already saved. Skipping re-run...")
  final_rf_fit <- readRDS(fit_path)
  final_rf_workflow <- readRDS(workflow_path)
  rf_predictions <- readRDS(pred_path)
  rf_metrics <- readRDS(metrics_path)

} else {
  message(" Finalizing and fitting Random Forest model...")

  # Extract best parameters
  best_params <- best_rf_params

  # Finalize the workflow
  final_rf_workflow <- finalize_workflow(rf_workflow, best_params)

  # Fit final model on training set
  set.seed(123)
  final_rf_fit <- fit(final_rf_workflow, data = train_data)

  # Predict on test set
  rf_predictions <- predict(final_rf_fit, new_data = test_data) %>%
    bind_cols(test_data %>% select(mortalities))

  # Evaluate performance
  rf_metrics <- rf_predictions %>%
    metrics(truth = mortalities, estimate = .pred)

  # Save all results
  saveRDS(final_rf_fit, file = fit_path)
  saveRDS(final_rf_workflow, file = workflow_path)
  saveRDS(rf_predictions, file = pred_path)
  saveRDS(rf_metrics, file = metrics_path)

  message("Final model fit and results saved.")
}

# Review metrics
print(rf_metrics)

```

The model is performing well, RMSE and MAE both reflect reasonable prediction error magnitudes, with reasonable error and strong explanatory power.

```{r}

# Load saved RF metrics
rf_metrics <- readRDS(here::here("results", "output", "rf_test_metrics.rds"))

# Convert to wide format for easy table summary
rf_summary_table <- rf_metrics %>%
  select(.metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  mutate(
    Model = "Random Forest",
    Notes = "Tuned with 4 predictors (mtry = 4), min_n = 15"
  ) %>%
  select(Model, RMSE = rmse, R2 = rsq, MAE = mae, Notes)

# Save the summary table
saveRDS(rf_summary_table, here::here("results", "tables", "model_performance_summary.rds"))

rf_summary_table
```

```{r}
## Random Forest: Predicted vs Observed Plot

# Load predictions (if not already in memory)
rf_predictions <- readRDS(here::here("results", "output", "rf_test_predictions.rds"))

# Create the predicted vs observed plot
rf_pred_plot <- ggplot(rf_predictions, aes(x = mortalities, y = .pred)) +
  geom_point(alpha = 0.4, color = "darkgreen") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Random Forest: Predicted vs Observed Mortality",
    x = "Observed Mortality",
    y = "Predicted Mortality"
  ) +
  theme_minimal()

# Save the plot to results/figures
ggsave(
  filename = here::here("results", "figures", "rf_pred_vs_obs.png"),
  plot = rf_pred_plot,
  width = 7, height = 5, dpi = 300)

rf_pred_plot
```

```{r}
## Random Forest: Variable Importance Plot

# Load fitted model (if not already in memory)
final_rf_fit <- readRDS(here::here("results", "output", "final_rf_fit.rds"))

# Create VIP plot
rf_vip_plot <- final_rf_fit %>%
  extract_fit_parsnip() %>%
  vip::vip(num_features = 10) +
  labs(title = "Top 10 Predictors (Random Forest)") +
  theme_minimal()

# Save the plot to results/figures
ggsave(
  filename = here::here("results", "figures", "rf_variable_importance.png"),
  plot = rf_vip_plot,
  width = 7)

rf_vip_plot
```

A Random Forest model was tuned and evaluated to predict county-level hypertension-related cardiovascular mortality. On the held-out test set, the model achieved an RMSE of 115.3, MAE of 66.9, and R² of 0.748, indicating that it explained approximately 75% of the variance in mortality rates across U.S. counties. These results reflect strong predictive performance and consistency with cross-validation metrics.

This model will serve as a benchmark against which we compare other algorithms, including regularized regression (e.g., LASSO).

####LASSO model

Define the LASSO model specification
```{r}
# LASSO uses glmnet engine with penalty tuning and fixed mixture = 1 (pure LASSO)
lasso_spec <- linear_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet") %>%
  set_mode("regression")

#Create the LASSO workflow
# Workflow with recipe
lasso_workflow <- workflow() %>%
  add_model(lasso_spec) %>%
  add_recipe(mortality_recipe)

```

Tune the penalty using cross-validation

```{r}
## Tune LASSO penalty parameter

# File paths
lasso_tuned_path <- here::here("results", "large-files", "lasso_tuned.rds")
lasso_metrics_path <- here::here("results", "output", "lasso_tuned_metrics.rds")
lasso_best_param_path <- here::here("results", "output", "lasso_best_params.rds")

# Load or tune
if (file.exists(lasso_tuned_path) && file.exists(lasso_metrics_path) && file.exists(lasso_best_param_path)) {
  message("Loading saved LASSO tuning results...")

  lasso_tuned <- readRDS(lasso_tuned_path)
  lasso_tuned_metrics <- readRDS(lasso_metrics_path)
  best_lasso_params <- readRDS(lasso_best_param_path)
} else {
  message("Running LASSO tuning...")

  lasso_grid <- grid_regular(penalty(range = c(-4, -1)), levels = 10)  # log10 scale

  set.seed(123)
  lasso_tuned <- tune_grid(
    lasso_workflow,
    resamples = cv_folds,
    grid = lasso_grid,
    metrics = metric_set(rmse, rsq, mae),
    control = control_grid(save_pred = TRUE))

  lasso_tuned_metrics <- collect_metrics(lasso_tuned)
  best_lasso_params <- select_best(lasso_tuned, metric = "rmse")

  # Save results
  saveRDS(lasso_tuned, lasso_tuned_path)
  saveRDS(lasso_tuned_metrics, lasso_metrics_path)
  saveRDS(best_lasso_params, lasso_best_param_path)

  message("LASSO Tuning complete and results saved.")
}

# Preview metrics
print(lasso_metrics)
```

Creeating a slimmed down version of the tuning RDS file to reduce file size for Github and working with if statements to load this from the correct folder and to not overwrite files.
```{r}
## ---- slim_lasso_model_save, message=TRUE, warning=FALSE -------------------

slim_lasso_path <- here::here("results", "output", "lasso_model_minimal.rds")

if (!file.exists(slim_lasso_path)) {
  message("Creating slimmed LASSO model object...")

  lasso_final <- extract_workflow(lasso_tuned)
  lasso_metrics <- lasso_tuned_metrics

  # Extract coefficients
  lasso_fit <- extract_fit_parsnip(lasso_final)
  lasso_coeffs <- as.matrix(lasso_fit$fit$beta)

  saveRDS(
    list(model = lasso_final, metrics = lasso_metrics, coefficients = lasso_coeffs),
    file = slim_lasso_path
  )

  message("Saved slimmed LASSO model as lasso_model_minimal.rds")
} else {
  message("Slimmed LASSO model already exists. Skipping save.")
}

```

The slim model can be toggled on or off here by changing from TRUE to FALSE if wanting to use the full version. The full RDS file is saved in the large-files folder which will not automatically upload to github.
```{r}
## ---- load_lasso_model, message=TRUE, warning=FALSE -------------------

use_slim_lasso_model <- TRUE  # Toggle to FALSE to use full model

if (use_slim_lasso_model) {
  if (!file.exists(here::here("results", "output", "lasso_model_minimal.rds"))) {
    stop("Slim LASSO model not found. Please run the slim-save chunk or disable slim mode.")
  }

  message("Loading slimmed LASSO model...")
  lasso_slim <- readRDS(here::here("results", "output", "lasso_model_minimal.rds"))

  lasso_final_model <- lasso_slim$model
  lasso_metrics <- lasso_slim$metrics
  lasso_coeffs <- lasso_slim$coefficients
  best_lasso_params <- readRDS(here::here("results", "output", "lasso_best_params.rds"))

} else {
  if (!file.exists(here::here("results", "large-files", "lasso_tuned.rds"))) {
    stop("Full LASSO model not found.")
  }

  message("Loading full LASSO model...")
  lasso_tuned <- readRDS(here::here("results", "large-files", "lasso_tuned.rds"))
  lasso_metrics <- readRDS(here::here("results", "output", "lasso_tuned_metrics.rds"))
  lasso_final_model <- extract_workflow(lasso_tuned)

  lasso_fit <- extract_fit_parsnip(lasso_final_model)
  lasso_coeffs <- as.matrix(lasso_fit$fit$beta)

  best_lasso_params <- readRDS(here::here("results", "output", "lasso_best_params.rds"))}

```

```{r}
#Plot LASSO Tuning Plot
# Plot RMSE vs penalty (log10 scale)
lasso_tune_plot <- lasso_metrics %>%
  filter(.metric == "rmse") %>%
  ggplot(aes(x = penalty, y = mean)) +
  geom_point(size = 2, color = "steelblue") +
  geom_line() +
  scale_x_log10() +
  theme_minimal() +
  labs(title = "LASSO Tuning Results (RMSE vs Penalty)",
       x = "Penalty (log10 scale)",
       y = "Mean RMSE")

# Save to results/figures
ggsave(filename = here::here("results", "figures", "lasso_tuning_plot.png"),
  plot = lasso_tune_plot,
  width = 7, height = 5, dpi = 300)

lasso_tune_plot
```

```{r}
## Final LASSO Model Fit and Evaluation

# Output paths
lasso_fit_path <- here::here("results", "output", "final_lasso_fit.rds")
lasso_workflow_path <- here::here("results", "output", "final_lasso_workflow.rds")
lasso_pred_path <- here::here("results", "output", "lasso_test_predictions.rds")
lasso_metrics_path <- here::here("results", "output", "lasso_test_metrics.rds")

if (file.exists(lasso_fit_path) &&
    file.exists(lasso_workflow_path) &&
    file.exists(lasso_pred_path) &&
    file.exists(lasso_metrics_path)) {

  message("Loading saved LASSO model and outputs...")
  final_lasso_fit <- readRDS(lasso_fit_path)
  final_lasso_workflow <- readRDS(lasso_workflow_path)
  lasso_test_predictions <- readRDS(lasso_pred_path)
  lasso_test_metrics <- readRDS(lasso_metrics_path)

} else {
  # Finalize model
  final_lasso_model <- finalize_model(lasso_spec, best_lasso_params)

  # Build workflow
  final_lasso_workflow <- workflow() %>%
    add_model(final_lasso_model) %>%
    add_recipe(mortality_recipe)

  # Fit and predict
  final_lasso_fit <- fit(final_lasso_workflow, data = train_data)

  lasso_test_predictions <- predict(final_lasso_fit, new_data = test_data) %>%
    bind_cols(test_data %>% select(mortalities))

  # Evaluate
  lasso_test_metrics <- lasso_test_predictions %>%
    metrics(truth = mortalities, estimate = .pred)

  # Save results
  saveRDS(final_lasso_fit, lasso_fit_path)
  saveRDS(final_lasso_workflow, lasso_workflow_path)
  saveRDS(lasso_test_predictions, lasso_pred_path)
  saveRDS(lasso_test_metrics, lasso_metrics_path)
}

# Print metrics
print(lasso_test_metrics)
```

```{r}
## LASSO: Predicted vs Observed Plot

# Load predictions (if not already in memory)
lasso_test_predictions <- readRDS(here::here("results", "output", "lasso_test_predictions.rds"))

lasso_pred_plot <- ggplot(lasso_test_predictions, aes(x = mortalities, y = .pred)) +
  geom_point(alpha = 0.4, color = "steelblue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "LASSO Regression: Predicted vs Observed Mortality",
    x = "Observed Mortality",
    y = "Predicted Mortality"
  ) +
  theme_minimal()

# Save to file
ggsave(
  filename = here::here("results", "figures", "lasso_pred_vs_obs.png"),
  plot = lasso_pred_plot,
  width = 7, height = 5, dpi = 300)

lasso_pred_plot
```

```{r}
lasso_summary_table <- lasso_test_metrics %>%
  select(.metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  mutate(
    Model = "LASSO Regression",
    Notes = paste("Penalty:", signif(best_lasso_params$penalty, 3))
  ) %>%
  select(Model, RMSE = rmse, R2 = rsq, MAE = mae, Notes)

saveRDS(lasso_summary_table, here::here("results", "tables", "lasso_model_summary.rds"))

lasso_summary_table
```

The LASSO regression model achieved a Root Mean Squared Error (RMSE) of ~119.63, with an R-squared of ~0.726, indicating a reasonably strong fit for predicting county-level hypertension-related mortality. Compared to the random forest model (RMSE ~115.30, R² ~0.748), LASSO performs slightly worse, but still retains good explanatory power.

Notably, LASSO applies regularization and performs variable selection, meaning it can identify the most influential predictors while shrinking the less important ones to zero. This makes it a useful, interpretable model for identifying risk factors — though it sacrifices some predictive accuracy compared to Random Forest.

Plot Predicted vs Observed Mortality

```{r}
ggplot(lasso_test_predictions, aes(x = mortalities, y = .pred)) +
  geom_point(alpha = 0.4, color = "steelblue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "LASSO Regression: Predicted vs Observed Mortality",
    x = "Observed Mortality",
    y = "Predicted Mortality"
  ) +
  theme_minimal(base_size = 13)

```

LASSO is a linear model with regularization — it struggles to capture high-order nonlinear patterns. The horizontal “bands” in predicted values suggest that LASSO may be underfitting in high or low mortality counties. While LASSO remains interpretable, random forest outperforms it in predictive power and flexibility, especially in a dataset with many categorical and potentially non-linear features.

###Comparing models

Generating a model comparison table:

```{r}

## Generate model comparison table with 95% confidence intervals

# Load test metrics
rf_metrics <- readRDS(here::here("results", "output", "rf_test_metrics.rds"))
lasso_metrics <- readRDS(here::here("results", "output", "lasso_test_metrics.rds"))

# Add model names
rf_metrics$model <- "Random Forest"
lasso_metrics$model <- "LASSO Regression"

# Combine
all_metrics <- bind_rows(rf_metrics, lasso_metrics)

# Check if std_err is available
has_se <- "std_err" %in% names(all_metrics)

# Define function to create CI string
format_ci <- function(mean, se = NA) {
  if (is.na(se) || se == 0) return(as.character(round(mean, 2)))
  lower <- round(mean - 1.96 * se, 2)
  upper <- round(mean + 1.96 * se, 2)
  paste0(round(mean, 2), " [", lower, ", ", upper, "]")
}

# Create final table
if (has_se) {
  # Use CI formatting
  all_metrics_ci <- all_metrics %>%
    select(model, .metric, mean = .estimate, std_err) %>%
    mutate(ci = mapply(format_ci, mean, std_err)) %>%
    select(model, .metric, ci) %>%
    pivot_wider(names_from = .metric, values_from = ci)
} else {
  # Fall back to just point estimates
  all_metrics_ci <- all_metrics %>%
    select(model, .metric, mean = .estimate) %>%
    mutate(ci = round(mean, 2)) %>%
    select(model, .metric, ci) %>%
    pivot_wider(names_from = .metric, values_from = ci)
}

# Build gt table
model_perf_ci_table <- all_metrics_ci %>%
  gt() %>%
  tab_header(
    title = "Model Performance Summary",
    subtitle = "Test Set Evaluation for Random Forest and LASSO"
  ) %>%
  cols_label(
    model = "Model",
    rmse = "RMSE",
    rsq = "R²",
    mae = "MAE"
  ) %>%
  fmt_markdown(columns = everything()) %>%
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12),
    column_labels.font.weight = "bold"
  )

# Save table
saveRDS(model_perf_ci_table, here::here("results", "tables", "model_performance_summary.rds"))

# View it
model_perf_ci_table

```

Additional comparison table without CIs:

```{r}
model_perf_numeric <- bind_rows(rf_metrics, lasso_metrics) %>%
  select(model, .metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate)

saveRDS(model_perf_numeric, here::here("results", "output", "model_performance_numeric.rds"))

model_perf_numeric
```

Aim 4 Summary
We compared two predictive models for county-level hypertension-related CVD mortality: a random forest model and a LASSO regression model. The random forest achieved the best performance, with an RMSE of 115.3 and R² of 0.748 on the test set. LASSO regression performed slightly worse (RMSE = 119.6, R² = 0.726) but offers the benefit of interpretability through feature selection. These results indicate strong overall predictive performance, with random forest serving as the best-performing model. Both models demonstrated solid predictive performance, with random forest performing best. The structured pattern seen in the predicted vs observed plots reflects the nature of the input data — with many counties having identical values for categorical groupings. The use of dummy variables, while necessary, introduces constraints that can create repeated predictions for common demographic combinations, especially in linear models like LASSO. This highlights a trade-off between model interpretability and flexibility.