### 2025-04-11 code/eda-code/eda.qmd

---
title: "CDC hypertension and CDV Mortality disease data from adults in the USA exploratory analysis script"
date: "2025-02-20"
output: html_document
---

# Setup

```{r}
#load needed packages. make sure they are installed.
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set saving/loading paths
library(knitr) #for nice tables and data integration
library(corrplot) #for correlation matrix visualization
library(sf) #for geospatial analysis
library(ggpubr) #for advanced visualization tools
library(ggplot2) #for plots
library(gt) #For a tidy table

# Install if needed: 
#install.packages("webshot")
# webshot::install_phantomjs()
```

Load the data.

```{r}
#Path to data. 
data_location <- here::here("data","processed-data","final_processed_data.rds")

#load data. 
final_processed_data <- readRDS(data_location)

# Extract key datasets
processed_data_no_percent <- final_processed_data[["processed_data_no_percent"]]
county_race_sex <- final_processed_data[["county_race_sex"]]
county_race <- final_processed_data[["county_race"]]
county_sex <- final_processed_data[["county_sex"]]
county_overall <- final_processed_data[["county_overall"]]

```

Running some EDA and checking summaries of data:

We are going to make use of loops and creating a list of the different datasets that we created to reduce separate code to run the descriptive statisitics for each dataset. We are not include the percent change dataset in these evaluations.

Creating this datset list:
```{r}
list_datasets <- list(
  "processed_data_no_percent" = processed_data_no_percent,
  "County Race-Sex" = county_race_sex,
  "County Race" = county_race,
  "County Sex" = county_sex,
  "County Overall" = county_overall)

```

We will run the descriptive statistics for each dataset (as a single loop)
```{r}
# Loop through each dataset and print basic statistics
summary_stats <- bind_rows(lapply(names(list_datasets), function(name) {
  stats <- list_datasets[[name]] %>%
    summarise(
      mean = mean(mortalities, na.rm = TRUE),
      median = median(mortalities, na.rm = TRUE),
      sd = sd(mortalities, na.rm = TRUE),
      min = min(mortalities, na.rm = TRUE),
      max = max(mortalities, na.rm = TRUE))
  stats$Dataset <- name  # Add dataset name for identification
  return(stats)}))

# Print summary statistics
tibble::as_tibble(summary_stats) %>% 
  relocate(Dataset) %>% 
  print()
```

# Identify potential missing values in each dataset
```{r}
cat("\n### Missing Data Summary:\n")
lapply(names(list_datasets), function(name) {
  missing_count <- sum(is.na(list_datasets[[name]]))
  cat(name, "has", missing_count, "missing values.\n")
})
```

We want to visualise the different datasets:

We will revisit the histogram with the percent change removed.
```{r}
# Visualize the distribution of mortality rates with a histogram
revised_mortalities_histogram <- ggplot(processed_data_no_percent, aes(x = mortalities)) +
  geom_histogram(binwidth = 20, fill = "blue", alpha = 0.7) +
  scale_x_continuous(breaks = seq(0, max(processed_data_no_percent$mortalities, na.rm = TRUE), by = 200)) +
  theme_minimal() +
  labs(title = "Distribution of Mortality Rates", x = "Mortality Rates", y = "Count")

saveRDS(revised_mortalities_histogram, file = here::here("data","processed-data", "revised_mortalities_histogram.rds"))
revised_mortalities_histogram <- readRDS(here::here("data","processed-data", "revised_mortalities_histogram.rds"))
# Print the mortalities histogram
print(revised_mortalities_histogram)
```

This is the total dataset and so the data is similar to the previous histogram. It is bimodal and there is a strong right skew and long tail. Peak mortalities are below 100, and we are not sure what accounts for the second peak.
This might be a factor of having two different age groups. We will see if this is made clear by splitting the age groups.

We want to split the data into two age groups one for individuals from 35 - 64 years, and the second for those 65 years and older. 
Age group 1 = 35 -64 years old
Age groups 2 = 65 years and older
```{r}
# Split into two age groups
age_group_1 <- processed_data_no_percent %>% filter(age_group == "Ages 35-64 years")
age_group_2 <- processed_data_no_percent %>% filter(age_group == "Ages 65+ years")

# Optional: Save these for later
saveRDS(age_group_1, here::here("data", "processed-data", "age_group_1.rds"))
saveRDS(age_group_2, here::here("data", "processed-data", "age_group_2.rds"))

```

We will recreate the histogram (one for each group) to check if we see distinct differences.

```{r}
# Histogram for Age Group 1 (35–64 years)
hist_age_group_1 <- ggplot(age_group_1, aes(x = mortalities)) +
  geom_histogram(fill = "steelblue", alpha = 0.7) +
  theme_minimal() +
  labs(title = "Distribution of Mortality Rates (Ages 35–64)",
    x = "Mortality Rate",
    y = "Count")

# Save new histo variable
saveRDS(hist_age_group_1, here::here("data", "processed-data", "hist_age_group_1.rds"))
#Print histogram
print(hist_age_group_1)

# Histogram for Age Group 2 (65+ years)
hist_age_group_2 <- ggplot(age_group_2, aes(x = mortalities)) +
  geom_histogram(binwidth = 20, fill = "darkorange", alpha = 0.7) +
  scale_x_continuous(breaks = seq(0, max(age_group_2$mortalities, na.rm = TRUE), by = 200)) +
  theme_minimal() +
  labs( title = "Distribution of Mortality Rates (Ages 65+)",
    x = "Mortality Rate",
    y = "Count")

# Save and optionally reload
saveRDS(hist_age_group_2, here::here("data", "processed-data", "hist_age_group_2.rds"))
# Print histogram 2
print(hist_age_group_2)

```

We will use a boxplot to get an overview of the other data:

```{r}
# Compare mortality rates across race-sex and sex stratifications

boxplot_race_sex <- ggplot(county_race_sex, aes(x = race_ethnicity, y = mortalities, fill = sex)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Mortality Rates by Race and Sex", x = "Race/Ethnicity", y = "Mortality Rate")

saveRDS(boxplot_race_sex, file = here::here("data","processed-data", "boxplot_race_sex.rds"))
print(boxplot_race_sex)

boxplot_sex <- ggplot(county_sex, aes(x = sex, y = mortalities)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Mortality Rates by Sex", x = "Sex", y = "Mortality Rate")

saveRDS(boxplot_sex, file = here::here("data","processed-data", "boxplot_sex.rds"))
print(boxplot_sex)

```

The box plots provide a good overview of the data nd indicate that there seem to be distinct trends and differences between races. 
Mortalities in the black and white race groups appear to be the highest, and there are many outliers in the white race group.
It also shows that there are differences between sexes but these are less marked and distinct than between races.
We will look at this in the statistical analysis.

Neatening up the boxplot for race without sex
```{r}
# Relationship between mortality and race
race_plot <- ggplot(county_race, aes(x = race_ethnicity, y = mortalities)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Mortality Rates by Race/Ethnicity", x = "Race/Ethnicity", y = "Mortality Rate") +
  theme(axis.text.x = element_text(size = 10)) +
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 10))

ggsave(here::here("results", "figures", "mortality_by_race.png"), plot = race_plot)
print(race_plot)

```
This is the same but without sex and the colours for a neater output.

We want an idea on any spatial trends in the data. We are using the sf function to plot the mortality data by the latitude and longitude.
```{r}
# Convert data to spatial format and create a map of mortality rates
# Ensure longitude and latitude values are numeric
processed_data_no_percent <- processed_data_no_percent %>%
  mutate(X_long = as.numeric(X_long), Y_lat = as.numeric(Y_lat))
# Convert to spatial object
sf_data <- st_as_sf(processed_data_no_percent, coords = c("X_long", "Y_lat"), crs = 4326)

# Generate a spatial plot of mortality rates
ggplot(sf_data) +
  geom_sf(aes(color = mortalities)) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Geospatial Distribution of Mortality Rates")

```

This is a very nice overview and can help us to evauate the trends further or to work on a more intricate plot. We must remember that we have not separated groups (especially age) yet.
But we do see that there is a lot more data from the East and SouthEast. Although it is possible that there might be limited data from some States on the West.

We want an overview of any temporal (time based) trends in the data. We will provide a proof of concept using a single county due to the overlap of data that may occur when plotting many different counties together. The statistical analysis will include further refinements. We are only working with the values where sex and race were both overall to reduce overlap.
```{r}
# Line plot of mortality rates over time
# Select a single county to illustrate the trend
selected_county <- "13037"  # County Location ID for Calhoun, GA
county_subset <- county_overall %>% filter(LocationID == selected_county)

temporal_trends_single_county <- ggplot(county_subset, aes(x = Year, y = mortalities, group = LocationID)) +
  geom_line(alpha = 0.6, color = "blue") +
  geom_smooth(se = FALSE, color = "red") +
  theme_minimal() +
  labs(title = paste("Mortality Rate Trends in Calhoun"), 
       x = "Year", y = "Mortality Rate")

print(temporal_trends_single_county)

```

This is a very strange plot and highlights the issues when combining different age groups together.
We will separate the age groups and run this again.

```{r}
# Use age-specific data
selected_county_data <- processed_data_no_percent %>%
  filter(LocationID == "13037",
         race_ethnicity == "Overall",
         sex == "Overall")

ggplot(selected_county_data, aes(x = Year, y = mortalities, color = age_group)) +
  geom_line(alpha = 0.6) +
  geom_smooth(se = FALSE) +
  theme_minimal() +
  labs(
    title = "Mortality Rate Trends by Age Group in Calhoun (Proof of Concept)",
    x = "Year", y = "Mortality Rate", color = "Age Group")

```

We see a clear increase in mortality rates in both age groups over time, but more so in the age group for 65+.

We plot the overall national trend over time with the two different age groups.
```{r}
# National mortality trend by age group
national_age_trend <- processed_data_no_percent %>%
  filter(race_ethnicity == "Overall", sex == "Overall") %>%
  group_by(Year, age_group) %>%
  summarise(mean_mortality = mean(mortalities, na.rm = TRUE)) %>%
  ggplot(aes(x = Year, y = mean_mortality, color = age_group)) +
  geom_line(size = 1) +
  geom_smooth(se = FALSE) +
  theme_minimal() +
  labs(title = "National Mortality Trends by Age Group",
       x = "Year", y = "Mean Mortality Rate", color = "Age Group")

saveRDS(national_age_trend, here::here("data", "processed-data", "national_trend_by_age.rds"))
print(national_age_trend)
```

```{r}
# National mortality trend by race
race_trend_65_plus <- processed_data_no_percent %>%
  filter(age_group == "Ages 65+ years", sex == "Overall") %>%
  group_by(Year, race_ethnicity) %>%
  summarise(mean_mortality = mean(mortalities, na.rm = TRUE)) %>%
  ggplot(aes(x = Year, y = mean_mortality, color = race_ethnicity)) +
  geom_line(size = 1) +
  geom_smooth(se = FALSE) +
  theme_minimal() +
  labs(title = "Mortality Trends by Race/Ethnicity (65+)",
       x = "Year", y = "Mean Mortality Rate", color = "Race")

saveRDS(race_trend_65_plus, here::here("data", "processed-data", "race_trend_65_plus.rds"))
print(race_trend_65_plus)

```

```{r}

race_trend_35_64 <- processed_data_no_percent %>%
  filter(age_group == "Ages 35-64 years", sex == "Overall") %>%
  group_by(Year, race_ethnicity) %>%
  summarise(mean_mortality = mean(mortalities, na.rm = TRUE)) %>%
  ggplot(aes(x = Year, y = mean_mortality, color = race_ethnicity)) +
  geom_line(size = 1) +
  theme_minimal() +
  labs(title = "Mortality Trends by Race/Ethnicity (Ages 35–64)",
       x = "Year", y = "Mean Mortality Rate", color = "Race")

saveRDS(race_trend_35_64, here::here("data", "processed-data", "race_trend_35_64.rds"))
print(race_trend_35_64)

```

```{r}
# National mortality trend by sex
sex_trend_65_plus <- processed_data_no_percent %>%
  filter(race_ethnicity == "Overall", age_group == "Ages 65+ years") %>%
  group_by(Year, sex) %>%
  summarise(mean_mortality = mean(mortalities, na.rm = TRUE)) %>%
  ggplot(aes(x = Year, y = mean_mortality, color = sex)) +
  geom_line(size = 1) +
  geom_smooth(se = FALSE) +
  theme_minimal() +
  labs(title = "Mortality Trends by Sex (65+)",
       x = "Year", y = "Mean Mortality Rate", color = "Sex")

saveRDS(sex_trend_65_plus, here::here("data", "processed-data", "sex_trend_65_plus.rds"))
print(sex_trend_65_plus)

```

```{r}

sex_trend_35_64 <- processed_data_no_percent %>%
  filter(age_group == "Ages 35-64 years", race_ethnicity == "Overall") %>%
  group_by(Year, sex) %>%
  summarise(mean_mortality = mean(mortalities, na.rm = TRUE)) %>%
  ggplot(aes(x = Year, y = mean_mortality, color = sex)) +
  geom_line(size = 1) +
  theme_minimal() +
  labs(title = "Mortality Trends by Sex (Ages 35–64)",
       x = "Year", y = "Mean Mortality Rate", color = "Sex")

saveRDS(sex_trend_35_64, here::here("data", "processed-data", "sex_trend_35_64.rds"))
print(sex_trend_35_64)

```

```{r}
# Grouped summary statistics
grouped_summary <- processed_data_no_percent %>%
  group_by(age_group, race_ethnicity, sex) %>%
  summarise( mean_mortality = round(mean(mortalities, na.rm = TRUE), 2),
    median_mortality = round(median(mortalities, na.rm = TRUE), 2),
    sd_mortality = round(sd(mortalities, na.rm = TRUE), 2),
    min_mortality = min(mortalities, na.rm = TRUE),
    max_mortality = max(mortalities, na.rm = TRUE),
    n = n(),.groups = "drop")

# Print summary as a nice table
grouped_summary %>% knitr::kable(caption = "Summary of Mortality by Age, Race, and Sex")

```

```{r}
# Define custom sorting levels
race_order <- c(
  "American Indian and Alaska Native",
  "Asian and Pacific Islander",
  "Black",
  "Hispanic",
  "White",
  "Overall")

sex_order <- c("Female", "Male", "Overall")

# Apply sorting and formatting
grouped_summary_clean <- grouped_summary %>%
  mutate(
     race_ethnicity = factor(race_ethnicity, levels = race_order),
    sex = factor(sex, levels = sex_order)
  ) %>%
  arrange(age_group, race_ethnicity, sex) %>%
  rename(
    `Age Group` = age_group,
    `Race/Ethnicity` = race_ethnicity,
    `Sex` = sex,
    `Mean Mortality` = mean_mortality,
    `Median Mortality` = median_mortality,
    `SD` = sd_mortality,
    `Minimum` = min_mortality,
    `Maximum` = max_mortality,
    `n` = n) %>%
  gt() %>%
  tab_header(
    title = "Summary of Mortality Rates by Age, Race, and Sex"
  ) %>%
  fmt_number(
    columns = c(`Mean Mortality`, `Median Mortality`, `SD`, `Minimum`, `Maximum`),
    decimals = 1
  ) %>%
  cols_align(align = "center") %>%
  tab_options(
    table.font.size = "small",
    heading.title.font.size = 14,
    data_row.padding = px(2))

gtsave(grouped_summary_clean, here::here("results", "tables", "summary_mortality_table.png"))

# View it
grouped_summary_clean

```

### Preliminary Linear Regression to Assess National Trends

```{r}
# This simple linear model estimates overall change in mortality over time.
# It helps confirm if a temporal trend exists before stratified modeling.

# Exploring mean mortality by year
yearly_means <- processed_data_no_percent %>%
  filter(race_ethnicity == "Overall", sex == "Overall") %>%
  group_by(Year) %>%
  summarise(mean_mortality = mean(mortalities, na.rm = TRUE))

ggplot(yearly_means, aes(x = Year, y = mean_mortality)) +
  geom_line() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Trend in Mean Mortality Rates Over Time ",
       x = "Year", y = "Mean Mortality Rate") +
  theme_minimal()

saveRDS(yearly_means, here::here("data", "processed-data", "yearly_means.rds"))
```

This model supports what the plots indicated that there are strong trends in the data indicating an increase in the mortality rates over time.

```{r}
eda_outputs <- list(
  processed_data_no_percent = processed_data_no_percent,
  age_group_1 = age_group_1,
  age_group_2 = age_group_2,
  grouped_summary = grouped_summary,
  county_race_sex = county_race_sex,
  county_race = county_race,
  county_sex = county_sex,
  county_overall = county_overall)

saveRDS(eda_outputs, here::here("data", "processed-data", "eda_outputs.rds"))

```

## Key Findings and Next Steps

- The distribution of mortality rates is right-skewed, suggesting potential transformations before modeling.
- Mortality rates vary markedly by race and sex, justifying stratified analysis.
- A proof-of-concept temporal trend was created for a single county (Calhoun, GA) using Location ID, ensuring unique selection.
- Temporal trend indicate a general increase the selected county, which should be modeled.
- All processed data has been saved into a single RDS file (`eda_outputs.rds`) for use in statistical analysis.

- Temporal and spatial variation in the data highlights the need for multilevel or mixed-effects modeling in the statistical analysis phase.
- The grouped summary by age, race, and sex provides key descriptive benchmarks and will support interpretation of model estimates.
- Notable outliers were observed in the White and Black race groups, which may warrant robustness checks or sensitivity analysis.