### 2025-04-11 products/manuscript/Manuscript.qmd

---
title: "Investigating the incidence and potential risk factors for hypertension-related mortalities in adults in the United States."
author: "Shaun van den Hurk"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
editor: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
#install.packages("magick")
library(here)
library(knitr)
library(magick)
```

**Authors**

-   First Autor$^{1,2,*}$ (ORCID: 0000-0000-1234-5678)
-   Second Author$^{3, *}$
-   Third Autor$^{3, \dagger}$
-   Last Author$^{1, 2, \land}$

**Author affiliations**

1.  College of Veterinary Medicine, University of Georgia, Athens, GA,
    USA.

$*$ These authors contributed equally to this work.

$\land$ Corresponding author: some\@email.com

$\dagger$ Disclaimer: The opinions expressed in this article are the
author's own and don't reflect their employer.

{{< pagebreak >}}

# Summary/Abstract

Lifestyle diseases such as cardiovascular disease make up the highest
causes of death in developed nations such as the United States. This
project focuses on hypertension-related cardiovascular disease and hopes
to help provide a greater understandign of the dynamics and patterns
that may exist with hypertension and cardiovascualr diseased related
mortality in the United States. The project seeks to evaluate data from
the CDC on the rates of hypertension-related cardiovascular disease
mortality and the associated demographic factors for this collected
data, including county, age group, race/ethnicity, and sex. The data
represents information collected from 2000 to 2019. This will be used to
see if there are any potential patterns or correlation between the rate
of mortality and these other demographic factors within the data.

{{< pagebreak >}}

# Introduction

## General Background Information

Lifestyle diseases such as cardiovascular disease make up the highest
causes of death in developed nations such as the United States. This
data analysis hopes to evaluate the rates of hypertension-related
cardiovascular disease mortality in adults in the United States and
trends that might be associated with demographic and geographic factors.

## Description of data and data source

The dataset was collected from the CDC website and the information was
sourced from the National Vital Statistics System. The dataset contains
rates and trends in the hypertension-related cardiovascular disease
(CVD) death rates. The study period is from 2000 to 2019, with two main
intervals displayed in the data 2000-2009, and 2010-2019. The data is
further distributed by county, age group (36-64 years and 65 years and
older), race/ethnicity, and sex.

## Questions/Hypotheses to be addressed

-   Hypertension-related cardiovascular disease mortalities are higher
    in select demographics (age groups, race/ethnicities and sex), and
    thus these demographics show correlated with mortalities rates.

-   There are clear regional/geographic distributions associated with
    hypertension-related cardiovascular disease mortalities.

Study Aims:
- Aim 1: Assess national temporal trends in hypertension-related CVD mortality, stratified by age group.
- Aim 2: Identify spatial clustering of county-level mortality.
- Aim 3: Evaluate disparities in mortality trends by race/ethnicity and sex, stratified by age group.
- Aim 4: Build predictive models using demographic and geographic features to estimate mortality trends.

Hypotheses:
- Mortality trends vary significantly over time and across age groups.
- County-level mortality exhibits spatial clustering, reflecting geographic disparities.
- Temporal trends differ significantly across race and sex, with widening inequities over time.
- Predictive models can identify key demographic and spatial factors associated with higher mortality trends.

{{< pagebreak >}}

# Methods

## Data import and cleaning

This data was downloaded from the CDC online database at the following
url:
https://data.cdc.gov/Heart-Disease-Stroke-Prevention/Rates-and-Trends-in-Hypertension-related-Cardiovas/uc9k-vc2j/about_data.
A csv file was downloaded which was almost 300MB in size. The initial
document/dataset contained 1103872 observations across 24 variables.
There were some initial processing steps performed to evaluate and
delete some of the unnecessary columns and to save this data to a .rds
document to reduce the size. This was all outlined in the
"processingcode" document, which will not run automatically because the
Excel (csv) file will not be available due to size concerns.

The intial exploration and cleaning and removal of the redundant columns
(columns with repeated single values without any additional
information). Some of the other columns were renamed to have the name of
the column reflect what these single value columns had contained
instead. The columns were checked to see what they contained and to see
how many unique entries there were and based upon this columns were
deleted and others renamed.

Deleted columns were: DataSource, Topic, Class,
StratificationCategory1/2/3, Data_Value_Footnote,
Data_Value_Footnote_Symbol, Geographic_Level, TopicID

Renamed columns: Stratification1 column to age_group, Stratification2
column to race_ethnicity, Stratification3 column to sex, Data_Value to
mortalities.

After the initial cleaning the data was evaluated again and some further
small cleaning and processing steps were performed (code is visible in
the processingfile-v1.qmd file) to reduce the number of rows by removing
NA values. The resulting dataframe contains 572572 observations and 14
variables.

Our summaries and the skimr function help to provide a good overview for
further evaluation of the data. We can see that our dataset contains a
combination of numeric and character data. The numeric data includes the
actual incidence rates of the hypertension-related cardiovascular
disease death rates, as well as the confidence intervals that were used,
and geographic coordinates. Our character data includes the race, sex,
age group, location data and the year.

This character data will be used as predictors of the mortality data as
we break down these different variables and look for potential patterns
and associations. The primary outcome of interest is the mortality rates
but the important potential predictors are the location, year,
race/ethnicity, and sex. AGe groups only provide two broad categories,
namely 35 to 65 years old and 65 years and older, which is a bit limited
but will allow for comparisons between these two groups.

Aim 1: We assessed national temporal trends in hypertension-related
cardiovascular disease (CVD) mortality from 2000 to 2019 using a series
of regression models. First, a linear regression model was fit to assess
the overall relationship between mortality rates and year. Next,
generalized linear models (GLMs) were used to evaluate time trends
stratified by race/ethnicity, sex, and age group. Interaction terms
(e.g., Year × race/ethnicity) were included to test for differential
trends across subgroups. Estimated annual changes in mortality rates
were compared across age groups using the emtrends and contrast
functions from the emmeans package.

Spatial Patterns of Mortality Aim 2: Spatial clustering analysis:
County-level mortality data were converted to spatial features using
geographic coordinates. Global and Local Moran’s I statistics were used
to assess spatial autocorrelation. Spatial weights were generated using
5-nearest-neighbor methods. Local clustering was visualized using
standardized mortality z-scores and LISA cluster classifications.

Evaluating National Demographic Disparities Aim 3: We fit generalized
linear models (GLMs) with interaction terms to estimate yearly changes
in hypertension-related CVD mortality rates from 2000–2019 across
demographic groups. Using emtrends() from the emmeans package, we
extracted slope estimates by race/ethnicity and sex, stratified by age
group (35–64 and 65+). Pairwise contrasts were conducted to evaluate
statistically significant differences in trends between groups.

Statistical Modeling (Aim 4) To predict county-level
hypertension-related cardiovascular disease (CVD) mortality, we trained
two supervised machine learning models: a random forest and a LASSO
regression. Models were trained on 80% of the data, with the remaining
20% held out for testing. Ten-fold cross-validation was used within the
training set for hyperparameter tuning. Predictors included year, age
group, race/ethnicity, and sex. The modeling dataset included demographic predictors (age group, race/ethnicity, sex) and time (year). Since several predictors were categorical, we applied dummy encoding (one-hot encoding) using step_dummy() within the recipes framework to convert these factors into binary indicator variables. We also normalized numeric predictors and removed any zero-variance predictors. Models were trained using a stratified 80/20 train/test split and evaluated using 10-fold cross-validation on the training set.
Random forest hyperparameters were selected from a regular grid of mtry and
min_n values, while LASSO regression used a log-scaled penalty grid.
Final model performance was evaluated using RMSE, MAE, and R² on the
test set. Confidence intervals were not computed for these metrics to
avoid additional computational burden.

## Statistical analysis

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

We began by exploring patterns in hypertension-related cardiovascular disease (CVD) mortality among U.S. adults aged 35 and older from 2000 to 2019. The EDA included summary statistics, histograms, and boxplots to assess mortality distributions overall and by key demographic subgroups (race/ethnicity, sex, age group). Temporal trends were visualized using line plots stratified by subgroup and geographic region. Spatial variation was explored with choropleth maps displaying county-level mortality rates across time and demographic strata. These descriptive analyses revealed substantial variation in mortality trends over time, with emerging disparities by race, sex, and age group. Additionally, spatial visualizations suggested clustering in certain states, laying the foundation for formal spatial and demographic analyses in later aims. All visualizations and cleaned summary tables were saved for downstream modeling.

```{r}
mortality_race_boxplot <- image_read(here::here("products", "manuscript", "supplement", "mortality_by_race.png"))
print(mortality_race_boxplot)

mortality_sex_boxplot <- image_read(here::here("products", "manuscript", "supplement", "mortality_by_sex.png"))
print(mortality_sex_boxplot)
```

## Basic statistical analysis

Spatial plots of mortalities per geographic regions were generated for
each race and compared on one plot. Two generalized linear models were
evaluated, the first examined how time, race/ethnicity, and sex (plot
below) influence mortality rates, and the second tested whether the
effect of time varied by race/ethnicity.

Overall, black individuals had the highest predcted mortality rates from
the generalized linear models, followed by white and native american
populations. Black individuals also had the highest hypertension-related
mortalities on the boxplot follwoed by white individuals. Females
appeared to have lower mortality rates than males across all races.

Race, sex and time all seem to be predictors of mortality. We see
differences in trends over time for different racial groups.

Our spatial clustering has revealed geographic clustering of
high-mortality regions, with high densities in the Eastern regions of
the country. There also appears to be a higher density of mortalities
from people in the race/ethnic group Black in certain regions. Asian and
Pacific Islander groups had relatively lower mortality rates and
geographic clustering.

```{r, echo=FALSE, message=FALSE}
#Print the glm plot
glm_interaction_plot <- image_read(here::here("products", "manuscript", "supplement", "glm_interaction_plot.png"))
print(glm_interaction_plot)
```

```{r, echo=FALSE, message=FALSE}
#Print the combined spatial plot
combined_spatial_plot <- image_read(here::here("products", "manuscript", "supplement", "combined_spatial_plot.png"))
print(combined_spatial_plot)
```

## Full analysis

Temporal Trends in Mortality Rates (2000–2019) (Aim 1): From 2000 to
2019, overall CVD mortality rates showed minimal change at the national
level. However, stratified models revealed substantial differences
across subgroups. Black individuals consistently exhibited the highest
mortality rates, though a declining trend was observed. In contrast,
rates among American Indian/Alaska Native individuals increased markedly
over time. Older adults (65+) had significantly higher mortality rates
and experienced a steeper upward trend compared to adults aged 35–64
(estimated slope: 1.22 vs. 1.04 deaths per 100,000 per year; p \<
0.001). These findings highlight persistent disparities and suggest
diverging subgroup trends in hypertension-related CVD mortality.

Spatial clustering analysis (Aim 2): We evaluated spatial patterns in
age-adjusted county-level hypertension-related cardiovascular disease
(CVD) mortality using global and local spatial autocorrelation methods.
Global spatial autocorrelation of county-level mortality rates was
statistically significant (Moran’s I = 0.0435, p \< 0.001), indicating a
weak but non-random spatial clustering pattern across the United States.
Local Indicators of Spatial Association (LISA) identified significant
clusters of high and low mortality values. LISA cluster types (e.g.,
high-high, low-low, spatial outliers) were visualized using
centroid-based maps for the contiguous U.S. and select states (Georgia,
Mississippi, Texas, Kentucky). Despite the overall geographic patterns,
only a small number of statistically significant local clusters were
detected, suggesting that while some regions exhibit elevated risk,
strong clustering at the county level may be limited or diffuse.

Investigating demographic disparities (Aim 3): Annual changes in
mortality rates varied substantially across demographic groups. Among
adults aged 35–64, Black individuals experienced the steepest declines,
while mortality increased most among males and American Indian/Alaska
Native populations. Similar patterns were observed among those aged 65+,
with significant disparities between groups. Pairwise comparisons
revealed the largest slope differences between Overall Male and Black
Overall groups, followed by disparities involving American Indian/Alaska
Native populations. These findings suggest widening inequities in
hypertension-related CVD mortality trends.

Predictive Modeling of County-Level Mortality (Aim 4) To evaluate the
predictive capacity of demographic and temporal factors on
hypertension-related cardiovascular mortality, we compared two machine
learning models: a random forest and a LASSO regression. Both models
were trained on 80% of the dataset and evaluated on a held-out 20% test
set, using cross-validation during tuning to optimize performance.

The random forest model demonstrated the strongest performance,
achieving a Root Mean Squared Error (RMSE) of 115.3, a Mean Absolute
Error (MAE) of 66.9, and an R² of 0.748 on the test set. In comparison,
the LASSO regression model yielded an RMSE of 119.6, MAE of 68.1, and R²
of 0.726. While slightly less accurate, the LASSO model offers valuable
interpretability by shrinking less important coefficients to zero,
effectively performing variable selection.

When comparing predicted vs observed plots, both models exhibited a distinct pattern of horizontal banding. This is likely due to the large number of categorical predictors (race, sex, age group) which were dummy-encoded. Many subsets of the population across counties share identical predictor combinations (e.g., "Black females aged 65+ in 2002"), leading to repeated predicted values for those groups. This results in stacked horizontal lines in the prediction plots, especially when using models like LASSO that produce linear fits with limited flexibility, and even in random forests when groupings dominate feature structure.

A comparison of model performance is summarized in Table (...X...).
These results suggest that both models capture meaningful patterns in
the data, but the random forest model more accurately predicts mortality
outcomes. Importantly, the use of interpretable models like LASSO can
aid in identifying potential risk factors despite modest reductions in
predictive power.

```{r, echo=FALSE, message=FALSE}
#Print the model performance summary
model_performance_summary_table <- readRDS(here::here("results", "tables", "model_performance_summary.rds"))
model_performance_summary_table
```

\_

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

Aim 3: These results show that hypertension-related CVD mortality trends
have not improved equally across demographic groups. While some
populations saw modest or no change, others—particularly Black
adults—experienced significant declines. Conversely, mortality rates
rose among males and American Indian/Alaska Native individuals. The
magnitude and significance of these differences reinforce the need for
targeted public health efforts to reduce disparities in cardiovascular
outcomes.

## Strengths and Limitations

## Conclusions

Note that this cited reference will show up at the end of the document,
the reference formatting is determined by the CSL file specified in the
YAML header. Many more style files for almost any journal [are
available](https://www.zotero.org/styles). You also specify the location
of your bibtex reference file in the YAML. You can call your reference
file anything you like.

{{< pagebreak >}}

# References